---
title: "Bison_Data_Filtering"
author: "Alex Depew"
date: "2026-05-25"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
# File information
This script is used to filter the bison microsatellite dataset used for further population genetics analyes.
 
Files used described below:
 
- `Bison_Genotypes_Raw(CORRECTED).csv`
   - Contains all samples for population genetic analyses
## Load Packages
```{r}
rm(list = ls())
library(tidyverse)
library(strataG)
library(adegenet)
library(PopGenReport)
```
 
## Import Datafiles
```{r}
raw_data <- read.csv("./datafiles/Bison_Genotypes_Raw(CORRECTED).csv",
                     check.names = FALSE)
 
# Cut blank columns
raw_data <- raw_data[ ,0:118]
```
 
## Herds of Interest
Change herd names for consistency 
```{r}
unique(raw_data$Herd)
names(raw_data)
 
herds_fixed <- raw_data %>% 
  mutate(Herd = case_when(
    Herd == "AB?" ~ "RL",
    Herd == "NORD/Liard" ~ "NORD",
    Herd == "Ronald Lake" ~ "RL", 
    Herd == "WBNPLakeOne" ~ "WBNPGR",
    Herd == "Mackenzie" ~ "MBS",
    Herd == "Yukon" ~ "AISH",
    Herd == "WBNPCarlson" ~ "WBNPHC",
    TRUE ~ Herd
)) %>% 
  # Some Delta Wood Bison are actually (WBNP Peace Athabasca DELTA)
  mutate(Herd = ifelse(
    Herd == "Delta" & Subspecies == "Wood Bison", 
    "WBNPPAD", 
    Herd
  ))
 
 
# Filter dataset to include wood bison herds ONLY and EINP Plains bison outgroup
target_herds <- c("EINPP",
                  "EINPW",
                  "WBNP",
                  "WBNPHC",
                  "WBNPGR",
                  "WBNPNY",
                  "WBNPPAD",
                  "WBNPLB",
                  "NORD",
                  "NAH",
                  "ETH",
                  "MBS",
                  "HZ",
                  "WENT",
                  "AISH",
                  "RL",
                  "WAB",
                  "CL")
 
wb_raw_herds <- herds_fixed %>% 
  filter(Herd %in% target_herds)
```
 
## Remove Loci 
```{r}
# Locus BM6017 and BMS911 are X-linked, and locus SE4748 is a sex marker
wb_locus_cleaned <- wb_raw_herds %>% 
  select(-BM6017, -BM6017_2, -BMS911, -BMS911_2, -SE4748, -SE4748_2)
```
 
## Remove Problem Samples
```{r}
wb_locus_cleaned[wb_locus_cleaned$Sample.ID == "Wab24_011", c("BMS601", "BMS601_2")] <- NA
```
 
 
## Duplicate Removal 
### Profile completeness
```{r}
# Many samples were collected in duplicate (ie. tissue and fecal)
 
# Assign a profile completeness metric to the sample to keep the highest covered sample
 
wb_locus_cleaned <- wb_locus_cleaned %>%
  # Convert "", "**", "0", "-99" to NA for all character columns
  mutate(across(
    where(is.character),
    ~ replace(.x, .x %in% c("", "**", "0", "-99"), NA)
  )) %>%
  # Calculates a row sum of non-NA values (% of profile amplified)
  mutate(COMPLETE = rowSums(!is.na(across(13:112)))) %>%
  # Moves newly created column with profile completeness (%) after AMPLIFICATION
  relocate(COMPLETE, .after = AMPLIFICATION)
 
# The "COMPLETE" column represents the coverage across the (now) 50 autosomal loci
```
 
### Remove low quality samples
```{r}
ggplot(wb_locus_cleaned, aes(x = COMPLETE)) +
  geom_bar(fill = "grey40", width = 1.5) +
  labs(
    x = "% Profile Amplified",
    y = "Number of Samples"
  ) +
  theme_bw(base_size = 12) +
  theme(
 
    axis.title = element_text(size = 10)
  )
 
# We will keep samples at a threshold of >=92% profile coverage of the remaining 50 autosomal markers
wb_locus_cleaned <- wb_locus_cleaned %>% 
  filter(COMPLETE >=92)
```
## Remove Duplicate Samples
### By ID
```{r}
which(duplicated(wb_locus_cleaned$Sample.ID)) # Many Sample IDs are duplicated

# Remove lowest quality duplicates
wb_cleaned_ids <- wb_locus_cleaned %>% 
  group_by(Sample.ID, Herd) %>% # Group duplicate IDs within herds together
  slice_max(order_by = COMPLETE, n = 1, with_ties = FALSE) %>% # Retain only the duplicate within a herd with the highest apparent genotype quality
  ungroup()
 

# Double check for duplicated IDs 
dupes <- wb_cleaned_ids %>% 
  group_by(Sample.ID, Herd) %>% 
  filter(n() > 1) %>% 
  ungroup()
 
dupes # This df contains the samples which have duplicate Sample.ID and are from the same Herd (some samples have duplicate IDs across herds and are therefore not at risk of being duplicates)
```
 
 
### By Genotype
```{r}
# Transform the dataframe into a genind then a gtypes object. First, combine locus columns
locus_cols <- names(wb_cleaned_ids)[14:ncol(wb_cleaned_ids)]
 
# Get base locus names without "_2"
base_loci <- unique(gsub("_2$", "", locus_cols))
 
# Loop through each locus and combine alleles
for (locus in base_loci) {
  col1 <- locus
  col2 <- paste0(locus, "_2")
  if (col1 %in% names(wb_cleaned_ids) && col2 %in% names(wb_cleaned_ids)) {
    wb_cleaned_ids[[col1]] <- paste(wb_cleaned_ids[[col1]], wb_cleaned_ids[[col2]], sep = "/")
  }
}
 
# Remove all the _2 columns
wb_cleaned_ids <- wb_cleaned_ids[, !grepl("_2$", names(wb_cleaned_ids))]
 
# Clean missing data values
wb_cleaned_ids[wb_cleaned_ids == "0/0"] <- NA
wb_cleaned_ids[wb_cleaned_ids == "**/**"] <- NA
wb_cleaned_ids[wb_cleaned_ids == "NA/NA"] <- NA
wb_cleaned_ids[wb_cleaned_ids == "-99/-99"] <- NA
 
# Import as a gtypes object with strataG package
wb.genind <- df2genind(X=wb_cleaned_ids[,c(14:ncol(wb_cleaned_ids))],
                              sep = "/",
                              ind.names = wb_cleaned_ids$Sample.ID,
                              loc.names = NULL,
                              pop = wb_cleaned_ids$Herd,
                              NA.char = "NA",
                              ploidy = 2,
                              type = "codom",
                              strata = NULL,
                              hierarchy = NULL)
 
wb.gtypes <- genind2gtypes(wb.genind)
```
 
### Check for duplicate profiles
Duplicate profiles will be identified if they share >% profile similarity
```{r}
# Make a df with the duplicates called with 90% or more profile similarity (ie. 45 or more exact loci)
dupe_profiles <- dupGenotypes(wb.gtypes, 
                  num.shared = 0.90)
 
```
 
### Remove duplicates from dataset
```{r}
# Pull 1 column of identified duplicates
remove_dupes <- dupe_profiles$ids.2
 
# Remove the duplicated IDs that were identified
wb_non_dup_ids_screened <- wb_cleaned_ids %>% 
  filter(!Sample.ID %in% remove_dupes)
```
 
# Missing Year Fix ----
```{r}
# Import list of genotypes used in Ball 2016 - 
ball_samples <- read.csv("./datafiles/ball_samples/Ball_2016_samples.csv", check.names = F)
 
ball_samples <- ball_samples %>% 
  # Fix the ID names to match what is present in the database
  mutate(
    IDrework = ID %>%
      gsub("^Bison-", "", .) %>%
      gsub("_", "-", .) %>% 
      gsub("RainbowLake1", "rainbow lake 1", .) %>% 
      gsub("Chinchaga-1", "chinchaga 1", .) %>% 
      gsub("Chinchaga-2", "chinchaga 2", .) %>% 
      gsub("RonaldLakeFetus", "ronald lake fetus", .)
  ) %>% 
  relocate(IDrework, .after = ID) %>% 
  # Make herd names consistent with the abbreviations in the main database
  mutate(
    HERD = case_when(
      Population == "RonaldLake"   ~ "RL",
      Population == "NWT"          ~ "MBS",
      Population == "Rainbow Lake" ~ "HZ",
      Population == "Wentzel"      ~ "WENT",
      Population == "Harpercreek"  ~ "WAB",
      Population == "Hay zama"     ~ "HZ",
      TRUE ~ Population
    )
  ) %>% 
  relocate(HERD, .after = Population)
```
 
 
 
## Fix Missing Collection Year
```{r}
year_fix <- wb_non_dup_ids_screened %>% 
  # From Ball 2016 paper
  mutate(
    Year.Collected = replace(
      Year.Collected,
      is.na(Year.Collected) &
      Sample.ID %in% ball_samples$IDrework &
      Herd %in% ball_samples$HERD,
      2015    
    )
  ,
    # Handle WR##### IDs
    Year.Collected = ifelse(
      is.na(Year.Collected) & grepl("^WR\\d{5}", Sample.ID),
      {
        two_digits <- str_sub(str_remove(Sample.ID, "^WR"), 1, 2)
        two_digits <- as.numeric(two_digits)
        ifelse(two_digits >= 50, 1900 + two_digits, 2000 + two_digits)
      },
      Year.Collected
    ),
    # Handle W##### IDs
    Year.Collected = ifelse(
      is.na(Year.Collected) & grepl("^W\\d{5}", Sample.ID),
      {
        two_digits <- str_sub(str_remove(Sample.ID, "^W"), 1, 2)
        two_digits <- as.numeric(two_digits)
        ifelse(two_digits >= 50, 1900 + two_digits, 2000 + two_digits)
      },
      Year.Collected
    ),
    # Add year collection for 25WBNP samples
    Year.Collected = ifelse(
      is.na(Year.Collected) & grepl("^25WBNP", Sample.ID),
      2025,
      Year.Collected
    ),
    # Add year collection for 2009 samples
    Year.Collected = ifelse(
      is.na(Year.Collected) & grepl("^2009", Sample.ID),
      2009,
      Year.Collected
    ),
    # Add year collection to specific samples based on 
    Year.Collected = case_when(
      Sample.ID == "YUK-223-Fe" ~ 2022,
      Sample.ID == "P86064" ~ 1986,
      Sample.ID == "P85033" ~ 1985,
      Sample.ID == "P85030" ~ 1985,
      Sample.ID == "P84030" ~ 1984,
      Sample.ID == "86063" ~ 1986,
      Sample.ID == "85029" ~ 1985,
      Sample.ID == "Tag3644" ~ 2010,
      Sample.ID == "BSN-61" ~ 2020,
      Sample.ID == "BSN-54" ~ 2020,
      Sample.ID == "BSN-53" ~ 2019,
      Sample.ID == "BSN-09" ~ 2016,
      Sample.ID == "BSN-06" ~ 2017,
      Sample.ID == "2196" ~ 2015,
      TRUE ~ Year.Collected
    ),
    # Add years from Ball et al. 2016
    Year.Collected = ifelse(
      is.na(Year.Collected) &
      Herd == "EINPW" &
      between(
        as.numeric(str_extract(Sample.ID, "\\d+")),
        88, 149
      ),
      2015,
      Year.Collected
    ),
    # Add years from Wilson 1999
    Year.Collected = ifelse(
      is.na(Year.Collected) &
      Herd == "EINPP" &
      between(
        as.numeric(str_extract(Sample.ID, "\\d+")),
        414, 478
      ),
      2015,
      Year.Collected
    ),
    Year.Collected = ifelse(
      is.na(Year.Collected) &
      grepl("^Mackenzie Bison Sanctuary", Sample.ID),
      1999,
      Year.Collected
    ),
      # Add year collection for Wentzel 2025 samples
    Year.Collected = ifelse(
      is.na(Year.Collected) & Herd == "WENT" & grepl("^BISO25", Sample.ID),
      2025,
      Year.Collected
    ),
    Year.Collected = ifelse(
      is.na(Year.Collected) & Herd == "EINPP" & grepl("2v", Sample.ID, ignore.case = TRUE),
      1990,
      Year.Collected
    ),
  # Add year collection to "Nov" and "Dec" labelled samples
  Year.Collected = ifelse(
   is.na(Year.Collected) & grepl("Nov", Sample.ID),
    2011,
    Year.Collected
),
   Year.Collected = ifelse(
  is.na(Year.Collected) & grepl("Dec", Sample.ID),
   2012,
  Year.Collected
),
 
  Year.Collected = ifelse(
    is.na(Year.Collected) & grepl("Etthithun 14-", Sample.ID),
    2014,
    Year.Collected
  ),
 
  Year.Collected = ifelse(
    is.na(Year.Collected) & grepl("Nordquist 14-", Sample.ID),
    2014,
    Year.Collected
  )
 
)
```
 
## Herd Assign
```{r}
herd_fix <- year_fix %>% 
  mutate(
    Herd = ifelse(
    grepl("WBNP_NY", Sample.ID),
    "WBNPNY",
    Herd
  ),
  Herd = ifelse(
    grepl("WBNP_GR", Sample.ID),
    "WBNPGR",
    Herd
  ),
  Herd = ifelse(
    grepl("WBNP_DL", Sample.ID),
    "WBNPPAD",
    Herd
  ),
  Herd = ifelse(
    grepl("WBNP_BF", Sample.ID),
    "WBNPLB",
    Herd
  ),
  Herd = ifelse(
    grepl("WBNP_19BF", Sample.ID),
    "WBNPLB",
    Herd
  ),
  Herd = ifelse(
    grepl("WBNP_20BF", Sample.ID),
    "WBNPLB",
    Herd
  ),
  Herd = ifelse(
    grepl("M.Grant Thesis", Year.Genotyped),
    "WBNPHC",
    Herd
  ),
  # Specific sample IDs 
  Herd = case_when(
    Sample.ID == "BSN82" ~ "WBNPHC",
    TRUE ~ Herd
      )
  )
```
 
```{r}
wb_filtered <- herd_fix %>% 
  # Remove samples without collection year metadata
  filter(
    !is.na(Year.Collected),
    !Herd == "WBNP")
```
 
# Format into GENIND Object
For allele frequencies / pop
```{r}
bison.genind <- df2genind(X=wb_filtered[,c(14:ncol(wb_filtered))],
                              sep = "/",
                              ind.names = wb_filtered$Sample.ID,
                              loc.names = NULL,
                              pop = wb_filtered$Herd,
                              NA.char = "NA",
                              ploidy = 2,
                              type = "codom",
                              strata = NULL,
                              hierarchy = NULL)
 
# Attach the Collection year to the genetic data
bison.genind@other$year <- wb_filtered$Year.Collected

temp1 <- bison.genind[bison.genind@other$year <= 2015]
temp2 <- bison.genind[bison.genind@other$year >= 2016]
 
# obtain population specific allele frequencies for whole dataset, temp1, and temp2
bison.genpop <- genind2genpop(bison.genind)
freqs <- as.data.frame(makefreq(bison.genpop))
write.csv(freqs, "./output/Whole_data_allele_frequencies2.csv", row.names = T)

temp1.genpop <- genind2genpop(temp1) # genind2genpop goes from individual data to population data
freqs1 <- as.data.frame(makefreq(temp1.genpop))
write.csv(freqs1, "./output/Temp1_allele_frequencies2.csv", row.names = T)

temp2.genpop <- genind2genpop(temp2) # genind2genpop goes from individual data to population data
freqs2 <- as.data.frame(makefreq(temp2.genpop))
write.csv(freqs2, "./output/Temp2_allele_frequencies2.csv", row.names = T)






# Allele frequency dataframe


freqs
# OR 
freqs2 <- allele.dist(bison.genind)
freqs2[[3]]

allele_freqs <- tab(temp1, NA.method = "mean")
 
```