---
title: "WBNP_Complex_Structure"
author: "Alex Depew"
date: "2026-08-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# File information
This script is used to evaluate the presence and extent of genetic substructure within the Wood Buffalo National Park complex. 

Files used described below:

- `Final_dataset_2026.06.28.csv`
   - Contains all samples for population genetic analyses, post-filtering
   
## Load Required Packages
```{r}
rm(list = ls())
library(adegenet)
library(graph4lg)
library(poppr)
library(hierfstat)
library(pophelper)
library(tidyverse)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(ggpubr)
library(grid)
```

## Load Required Data
```{r}
cleaned.data <- read.csv("./datafiles/Final_dataset_2026.07.17.csv",
                         check.names = F)
```

#### Convert data to Genind object
```{r}
# Make genind object
bison.genind <- df2genind(X=cleaned.data[,c(15:ncol(cleaned.data))],
                              sep = "/",
                              ind.names = cleaned.data$Spreadsheet_ID,
                              loc.names = NULL,
                              pop = cleaned.data$Herd,
                              NA.char = "NA",
                              ploidy = 2,
                              type = "codom",
                              strata = NULL,
                              hierarchy = NULL)

# Attach collection year metadata
bison.genind@other$year <- cleaned.data$Year.Collected
```

#### Subset data 
Data is subset into two temporal groups. Temporal groups are: 1: pre-2016, and 3: post-2016
```{r}
# Create datasets containing samples of interest for temporal comparisons
c1 <- bison.genind[bison.genind@pop %in% c("WBNPGR", "WBNPHC", 
                                           "WBNPLB", "WBNPNY", 
                                           "WBNPPAD", "WENT", 
                                           "RL", "WAB", "MBS") &
                   bison.genind@other$year <=2015,]

c2 <- bison.genind[bison.genind@pop %in% c("WBNPGR", "WBNPHC", 
                                           "WBNPLB", "WBNPNY", 
                                           "WBNPPAD", "WENT", 
                                           "RL", "WAB", "MBS") &
                   bison.genind@other$year >=2016,]
```

# Principal Coordinate Analysis (PCA)
Visualize population structure without assumptions of HWE or LD, to identify clusters in the data.
```{r}
herds <- c( "WBNPLB", "WBNPNY", "WBNPHC", "WBNPPAD","WBNPGR", "WENT", "RL", "WAB", "MBS")

herd_palette <- c("#CCBB44","#228833", "#4477AA", "#EE6677",  "#66CCEE", "#AA3377", "#BBBBBB", "#EE7733", "#332288")

herd_cols <- setNames(as.character(herd_palette), herds)

```

# PCA
### Perform PCA on both datasets
```{r}
# Perform PCA
x.c1 <- tab(c1, freq = TRUE, NA.method = "mean")
pca.c1 <- dudi.pca(x.c1, center = T, scale = F, scannf = F, nf = 2)

# Axes % Variance
eig.perc.c1 <- 100*pca.c1$eig/sum(pca.c1$eig)

x.c2 <- tab(c2, freq = TRUE, NA.method = "mean")
pca.c2 <- dudi.pca(x.c2, center = T, scale = F, scannf = F, nf = 2)

# Axes % Variance
eig.perc.c2 <- 100*pca.c2$eig/sum(pca.c2$eig)

#### axis limits
xlims <- c(-2.5, 3.5)
ylims <- c(-2, 2.1)
```

## PCA plot with ggplot
#### WBNP Complex Temp1
```{r}
c1@pop <- factor(c1@pop, levels = sort(unique(as.character(c1@pop))))
levels_c1 <- levels(c1@pop)
c1_cols <- herd_cols[levels_c1]

pca.c1$li[ ,1] <- pca.c1$li[ ,1] * -1

# Plot PCA object
pcac1 <- fviz_pca_ind(pca.c1,
             axes = c(1, 2),
             geom = "blank",
             habillage = c1@pop,
             repel = TRUE,
             palette = c1_cols,
             alpha.ind = 1,) + 
  stat_ellipse(aes(color = c1@pop), type = "norm", level = 0.68) + 
  geom_point(aes(colour = c1@pop),
             shape = 19,
             alpha = 0.75,
             size = 2) + 
#coord_fixed(ratio = 1, xlim = xlims, ylim = ylims) +
  
theme(
    plot.title = element_blank(),
    legend.title = element_text(face = "bold", size = 15),
    legend.text = element_text(size = 15),
    axis.title = element_text(face = "bold", size = 20),
    text = element_text(size = 20),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),  ) + 


  labs(title = NULL,
       # Fixed the placement of closing parentheses and commas below
       x = paste0("PC1 (", round(eig.perc.c1[1], 2), "%)"),
       y = paste0("PC2 (", round(eig.perc.c1[2], 2), "%)"),
       colour = "Herd") + 
    guides(colour = guide_legend(override.aes = list(size = 4, shape = 19)))

       

pcac1

ggsave("./output/plots/PCAs/wbnp_complex//temp1.svg",
       plot = pcac1, 
       device = "svg", 
       width = 12, 
       height = 12, 
       units = "in")
```
#### WBNP Complex Temp2
```{r}
c2@pop <- factor(c2@pop, levels = sort(unique(as.character(c2@pop))))
levels_c2 <- levels(c2@pop)
c2_cols <- herd_cols[levels_c2]

pca.c2$li[ ,1] <- pca.c2$li[ ,1] * -1
pca.c2$li[ ,2] <- pca.c2$li[ ,2] * -1


# Plot PCA object
pcac2 <- fviz_pca_ind(pca.c2,
             axes = c(1, 2),
             geom = "blank",
             habillage = c2@pop,
             repel = TRUE,
             palette = c2_cols,
             alpha.ind = 1,) + 
  stat_ellipse(aes(color = c2@pop), type = "norm", level = 0.68) + 
  geom_point(aes(colour = c2@pop),
             shape = 19,
             alpha = 0.75,
             size = 2) + 
#coord_fixed(ratio = 1, xlim = xlims, ylim = ylims) +
  
theme(
    plot.title = element_blank(),
    legend.title = element_text(face = "bold", size = 15),
    legend.text = element_text(size = 15),
    axis.title = element_text(face = "bold", size = 20),
    text = element_text(size = 20),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),  ) + 


  labs(title = NULL,
       # Fixed the placement of closing parentheses and commas below
       x = paste0("PC1 (", round(eig.perc.c2[1], 2), "%)"),
       y = paste0("PC2 (", round(eig.perc.c2[2], 2), "%)"),
       colour = "Herd") + 
    guides(colour = guide_legend(override.aes = list(size = 4, shape = 19)))

       

pcac2

ggsave("./output/plots/PCAs/wbnp_complex//temp2.svg",
       plot = pcac2, 
       device = "svg", 
       width = 12, 
       height = 12, 
       units = "in")
```
# Create STRUCTURE Files
```{r}
summary(c1)
# W1 structure (202 individuals, 9 pops)
c1.str <- c1 
pop(c1.str) <- as.numeric(as.factor(pop(c1.str)))

genind_to_structure(c1.str,
                    output = "STRUCTURE/wbnp_complex/c1.structure.txt")

summary(c2)
# W2 structure (269 individuals, 9 pops)
c2.str <- c2
pop(c2.str) <- as.numeric(as.factor(pop(c2.str)))

genind_to_structure(c2.str,
                    output = "STRUCTURE/wbnp_complex/c2.structure.txt")
```

