---
title: "WBNP_PCA_STRUCT"
author: "Alex Depew"
date: "2026-07-03"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# File information
This script is used to evaluate the presence and extent of genetic substructure within Wood Buffalo National Park. 

Files used described below:

- `Final_dataset_2026.06.28.csv`
   - Contains all samples for population genetic analyses, post-filtering
   
## Load Required Packages
```{r}
rm(list = ls())
library(adegenet)
library(graph4lg)
library(poppr)
library(hierfstat)
library(pophelper)
library(tidyverse)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(ggpubr)
library(grid)
```

## Load Required Data
```{r}
cleaned.data <- read.csv("./datafiles/Final_dataset_2026.07.17.csv",
                         check.names = F)
```

#### Convert data to Genind object
```{r}
# Make genind object
bison.genind <- df2genind(X=cleaned.data[,c(15:ncol(cleaned.data))],
                              sep = "/",
                              ind.names = cleaned.data$Spreadsheet_ID,
                              loc.names = NULL,
                              pop = cleaned.data$Herd,
                              NA.char = "NA",
                              ploidy = 2,
                              type = "codom",
                              strata = NULL,
                              hierarchy = NULL)

# Attach collection year metadata
bison.genind@other$year <- cleaned.data$Year.Collected
```

#### Subset data 
Data is subset into two temporal groups. Temporal groups are: 1: pre-2016, and 3: post-2016
```{r}
# Create datasets containing samples of interest for temporal comparisons
w1 <- bison.genind[bison.genind@pop %in% c("WBNPGR", "WBNPHC", 
                                           "WBNPLB", "WBNPNY", "WBNPPAD", "WENT") &
                   bison.genind@other$year <=2015,]

w2 <- bison.genind[bison.genind@pop %in% c("WBNPGR", "WBNPHC", 
                                           "WBNPLB", "WBNPNY", "WBNPPAD", "WENT") &
                   bison.genind@other$year >=2016,]
```

# Principal Coordinate Analysis (PCA)
Visualize population structure without assumptions of HWE or LD, to identify clusters in the data.
```{r}
wbnp_subherds <- c("WBNPPAD", "WENT", "WBNPGR", "WBNPHC", "WBNPNY", "WBNPLB")

wbnp_palette <- c("#cecece", 
                  "#a559aa", 
                  "#59a89c", 
                  "#f0c571", 
                  "#e02b35", 
                  "#082a54"
                  )

wbnp_palette <- c("#d9ed92", "#99d98c", "#52b69a", "#168aad", "#1a759f", "#184e77")


wbnp_palette <-  c("#4477AA", "#EE6677",  "#66CCEE", "#AA3377", "#BBBBBB", "#EE7733")


wbnp_cols <- setNames(as.character(wbnp_palette), wbnp_subherds)

# STRUCTURE colours
master_cols <- c("#ee964c", "#3b96cf", "#7d2a7c", "#e39bc1", "#e73f4e", 
                 "#4b9b46", "#f5df93", "#858434", "#ba552b", "#81c7c7", 
                 "#2d4a5e", "#44c29a", "#542e44", "#ff7f32", "#607d8b", "#1b4332")

```

# PCA
### Perform PCA on both datasets
```{r}
# Perform PCA
x.w1 <- tab(w1, freq = TRUE, NA.method = "mean")
pca.w1 <- dudi.pca(x.w1, center = T, scale = F, scannf = F, nf = 2)

# Axes % Variance
eig.perc.w1 <- 100*pca.w1$eig/sum(pca.w1$eig)

x.w2 <- tab(w2, freq = TRUE, NA.method = "mean")
pca.w2 <- dudi.pca(x.w2, center = T, scale = F, scannf = F, nf = 2)

# Axes % Variance
eig.perc.w2 <- 100*pca.w2$eig/sum(pca.w2$eig)

#### axis limits
xlims <- c(-2.5, 3.5)
ylims <- c(-2, 2.1)
```

## PCA plot with ggplot
#### WBNP Temp1 (W1: No EINP)
```{r}
w1@pop <- factor(w1@pop, levels = sort(unique(as.character(w1@pop))))
levels_w1 <- levels(w1@pop)
w1_cols <- wbnp_cols[levels_w1]


# Plot PCA object
pcaw1 <- fviz_pca_ind(pca.w1,
             axes = c(1, 2),
             geom = "blank",
             habillage = w1@pop,
             repel = TRUE,
             palette = w1_cols,
             alpha.ind = 1,) + 
  stat_ellipse(aes(colour = Groups), type = "norm", level = 0.68) + 
  geom_point(aes(colour = Groups),
             shape = 19,
             alpha = 0.75,
             size = 2) + 
coord_fixed(ratio = 1, xlim = xlims, ylim = ylims) +
  
theme(
    plot.title = element_blank(),
    legend.title = element_text(face = "bold", size = 15),
    legend.text = element_text(size = 15),
    axis.title = element_text(face = "bold", size = 20),
    text = element_text(size = 20),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),  ) + 

  labs(title = NULL,
       # Fixed the placement of closing parentheses and commas below
       x = paste0("PC1 (", round(eig.perc.w1[1], 2), "%)"),
       y = paste0("PC2 (", round(eig.perc.w1[2], 2), "%)"),
       colour = "Herd") + 
    guides(colour = guide_legend(override.aes = list(size = 4, shape = 19)))

pcaw1

ggsave("./output/plots/PCAs/wbnp/temp1_wbnp.svg",
       plot = pcaw1, 
       device = "svg", 
       width = 12, 
       height = 12, 
       units = "in")
```




```{r}
w2@pop <- factor(w2@pop, levels = sort(unique(as.character(w2@pop))))
levels_w2 <- levels(w2@pop)
w2_cols <- wbnp_cols[levels_w2]

# Plot PCA object
pcaw2 <- fviz_pca_ind(pca.w2,
             axes = c(1, 2),
             geom = "blank",
             habillage = w2@pop,
             repel = TRUE,
             palette = w2_cols,
             alpha.ind = 1) + 
  stat_ellipse(aes(color = w2@pop), type = "norm", level = 0.68) + 
  geom_point(aes(colour = w2@pop),
             shape = 19,
             alpha = 0.75,
             size = 2) + 
  coord_fixed(ratio = 1, xlim = xlims, ylim = ylims) +
  
theme(
    plot.title = element_blank(),
    legend.title = element_text(face = "bold", size = 15),
    legend.text = element_text(size = 15),
    axis.title = element_text(face = "bold", size = 20),
    text = element_text(size = 20),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),  ) + 


  labs(title = NULL,
       x = paste0("PC1 (", round(eig.perc.w2[1], 2), "%)"),
       y = paste0("PC2 (", round(eig.perc.w2[2], 2), "%)"),
       colour = "Herd") + 
    guides(colour = guide_legend(override.aes = list(size = 4, shape = 19)))

pcaw2

ggsave("./output/plots/PCAs/wbnp/temp2_wbnp.svg",
       plot = pcaw2, 
       device = "svg", 
       width = 12, 
       height = 12, 
       units = "in")
```

### Arrange Plots together
```{r}
#pcaw1_nl <- pcaw1 + theme(legend.position = "none") +
#  labs(tag = "C") +
#   theme(
#    plot.tag.position = "topleft",
#    plot.tag = element_text(face = "bold", size = 25),
#    plot.margin = margin(t = -25, r = 5, b = 0, l = 5, unit = "pt"))
    
#pcaw2_nl <- pcaw2 + theme(legend.position = "none") +
#  labs(tag = "D") +
#   theme(
#    plot.tag.position = "topleft",
#    plot.tag = element_text(face = "bold", size = 25),
#    plot.margin = margin(t = -25, r = 5, b = 0, l = 5, unit = "pt"))

#legendw1 <- get_legend(pcaw1)

#pcaw1: Keep legend and place it inside the plot area

```
## Add EINPW AND EINPP
#### Subset data 
Data is subset into two temporal groups. Temporal groups are: 1: pre-2016, and 3: post-2016
```{r}
# Create datasets containing samples of interest for temporal comparisons
a1 <- bison.genind[bison.genind@pop %in% c("EINPP", "EINPW", "WBNPGR", "WBNPHC", 
                                           "WBNPLB", "WBNPNY", "WBNPPAD", "WENT") &
                   bison.genind@other$year <=2015,]

a2 <- bison.genind[bison.genind@pop %in% c("EINPP", "EINPW", "WBNPGR", "WBNPHC", 
                                           "WBNPLB", "WBNPNY", "WBNPPAD", "WENT") &
                   bison.genind@other$year >=2016,]
```

# Principal Coordinate Analysis (PCA)
Visualize population structure without assumptions of HWE or LD, to identify clusters in the data.
```{r}
admix_subherds <- c("EINPP", "EINPW", "WBNPPAD", "WENT", "WBNPGR", "WBNPHC", "WBNPNY", "WBNPLB")



admix_palette <- c("#CCBB44","#228833", "#4477AA", "#EE6677",  "#66CCEE", "#AA3377", "#BBBBBB", "#EE7733")

admix_cols <- setNames(as.character(admix_palette), admix_subherds)
```

# PCA
### Perform PCA on both datasets
```{r}
# Perform PCA
x.a1 <- tab(a1, freq = TRUE, NA.method = "mean")
pca.a1 <- dudi.pca(x.a1, center = T, scale = F, scannf = F, nf = 2)

# Axes % Variance
eig.perc.a1 <- 100*pca.a1$eig/sum(pca.a1$eig)

x.a2 <- tab(a2, freq = TRUE, NA.method = "mean")
pca.a2 <- dudi.pca(x.a2, center = T, scale = F, scannf = F, nf = 2)

# Axes % Variance
eig.perc.a2 <- 100*pca.a2$eig/sum(pca.a2$eig)

#### axis limits
xlims <- c(-2.5, 3.5)
ylims <- c(-2, 2.1)
```

## PCA plot with ggplot
#### WBNP Temp1 (a1: With EINP)
```{r}
a1@pop <- factor(a1@pop, levels = sort(unique(as.character(a1@pop))))
levels_a1 <- levels(a1@pop)
a1_cols <- admix_cols[levels_a1]

#pca.a1$li[ ,2] <- pca.a1$li[ ,2] * -1

# Plot PCA object
pcaa1 <- fviz_pca_ind(pca.a1,
             axes = c(1, 2),
             geom = "blank",
             habillage = a1@pop,
             repel = TRUE,
             palette = a1_cols,
             alpha.ind = 1,) + 
  stat_ellipse(aes(color = a1@pop), type = "norm", level = 0.68) + 
  geom_point(aes(colour = a1@pop),
             shape = 19,
             alpha = 0.75,
             size = 2) + 
coord_fixed(ratio = 1, xlim = xlims, ylim = ylims) +
  
theme(
    plot.title = element_blank(),
    legend.title = element_text(face = "bold", size = 15),
    legend.text = element_text(size = 15),
    axis.title = element_text(face = "bold", size = 20),
    text = element_text(size = 20),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),  ) + 


  labs(title = NULL,
       # Fixed the placement of closing parentheses and commas below
       x = paste0("PC1 (", round(eig.perc.a1[1], 2), "%)"),
       y = paste0("PC2 (", round(eig.perc.a1[2], 2), "%)"),
       colour = "Herd") + 
    guides(colour = guide_legend(override.aes = list(size = 4, shape = 19)))

       

pcaa1

ggsave("./output/plots/PCAs/wbnp/temp1_admix.svg",
       plot = pcaa1, 
       device = "svg", 
       width = 12, 
       height = 12, 
       units = "in")
```


## PCA plot with ggplot
#### WBNP Temp1 (a2: With EINP)
```{r}
a2@pop <- factor(a2@pop, levels = sort(unique(as.character(a2@pop))))
levels_a2 <- levels(a2@pop)
a2_cols <- admix_cols[levels_a2]

pca.a2$li[ ,1] <- pca.a2$li[ ,1] * -1

# Plot PCA object
pcaa2 <- fviz_pca_ind(pca.a2,
             axes = c(1, 2),
             geom = "blank",
             habillage = a2@pop,
             repel = TRUE,
             palette = a2_cols,
             alpha.ind = 1,) + 
  stat_ellipse(aes(color = a2@pop), type = "norm", level = 0.68) + 
  geom_point(aes(colour = a2@pop),
             shape = 19,
             alpha = 0.75,
             size = 2) + 
coord_fixed(ratio = 1, xlim = xlims, ylim = ylims) +
  
theme(
    plot.title = element_blank(),
    legend.title = element_text(face = "bold", size = 15),
    legend.text = element_text(size = 15),
    axis.title = element_text(face = "bold", size = 20),
    text = element_text(size = 20),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),  ) + 


  labs(title = NULL,
       # Fixed the placement of closing parentheses and commas below
       x = paste0("PC1 (", round(eig.perc.a2[1], 2), "%)"),
       y = paste0("PC2 (", round(eig.perc.a2[2], 2), "%)"),
       colour = "Herd") + 
    guides(colour = guide_legend(override.aes = list(size = 4, shape = 19)))

       

pcaa2

ggsave("./output/plots/PCAs/wbnp/temp2_admix.svg",
       plot = pcaa2, 
       device = "svg", 
       width = 16, 
       height = 16, 
       units = "in")
```

### Arrange Plots together
```{r}
pcaa1_nl <- pcaa1 + theme(legend.position = "none") +
  labs(tag = "C") +
   theme(
    plot.tag.position = "topleft",
    plot.tag = element_text(face = "bold", size = 25)
  )
pcaa2_nl <- pcaa2 + theme(legend.position = "none") +
  labs(tag = "D") +
   theme(
    plot.tag.position = "topleft",
    plot.tag = element_text(face = "bold", size = 25)
  )

legenda1 <- get_legend(pcaa1)

admix_pcas <- grid.arrange(pcaa1_nl, legenda1, pcaa2_nl, # Set out grid
                            ncol = 3, 
                            widths = c(10, 2.5, 10)
                            )
                            
# Save plot
library(svglite)
ggsave("./output/plots/PCAs/wbnp/wbnp_admix_pcas.svg",
       plot = admix_pcas, 
       device = "svg", 
       width = 16, 
       height = 12, 
       units = "in")
```
# STRUCTURE Analysis
```{r}
# W1 structure (92 individuals, 6 pops)
w1.str <- w1 
pop(w1.str) <- as.numeric(as.factor(pop(w1.str)))

genind_to_structure(w1.str,
                    output = "STRUCTURE/wbnp/no_einp/w1.structure.txt")

# W2 structure (206 individuals, 6 pops)
w2.str <- w2
pop(w2.str) <- as.numeric(as.factor(pop(w2.str)))

genind_to_structure(w2.str,
                    output = "STRUCTURE/wbnp/no_einp/w2.structure.txt")

# A1 structure (233 individuals, 8 pops)
a1.str <- a1
pop(a1.str) <- as.numeric(as.factor(pop(a1.str)))

genind_to_structure(a1.str,
                    output = "STRUCTURE/wbnp/w_einp/a1.structure.txt")

# A2 structure (263 individuals, 8 pops)
a2.str <- a2
pop(a2.str) <- as.numeric(as.factor(pop(a2.str)))

genind_to_structure(a2.str,
                    output = "STRUCTURE/wbnp/w_einp/a2.structure.txt")

```
## Load Files
```{r}
# List files
# Full datasets
sfiles_w1 <- list.files("./STRUCTURE/wbnp/no_einp/results/temp1/", full.names = TRUE)
sfiles_w2 <- list.files("./STRUCTURE/wbnp/no_einp/results/temp2/", full.names = TRUE)

# Qlist (pophelper)
slist_w1 <- readQ(files = sfiles_w1, filetype = "structure")
slist_w2 <- readQ(files = sfiles_w2, filetype = "structure")

# Summarise runs
struct_sum_w1 <- summariseQ(tabulateQ(slist_w1))
struct_sum_w2 <- summariseQ(tabulateQ(slist_w2))

# And for admixed -----
# List files
# Full datasets
sfiles_a1 <- list.files("./STRUCTURE/wbnp/w_einp/results/temp1", full.names = TRUE)
sfiles_a2 <- list.files("./STRUCTURE/wbnp/w_einp/results/temp2/", full.names = TRUE)

# Qlist (pophelper)
slist_a1 <- readQ(files = sfiles_a1, filetype = "structure")
slist_a2 <- readQ(files = sfiles_a2, filetype = "structure")

# Summarise runs
struct_sum_a1 <- summariseQ(tabulateQ(slist_a1))
struct_sum_a2 <- summariseQ(tabulateQ(slist_a2))
```


## Likelihood Plots
```{r}
## WBNP1 ----
# K plots
w1.k <- evannoMethodStructure(data = struct_sum_w1, exportplot = F,
                           returnplot = T, returndata = F,
                           basesize = 25, linesize = 0.7,
                           height = 25, width = 25, units = "cm",
                           outputfilename = "w1_K",
                           exportpath = "./STRUCTURE_files/",
                           imgtype = "pdf")

p_lik <- w1.k$grobs[[1]]
p_dk <- w1.k$grobs[[4]]


# Label Likelihood as 'a'
panel_a <- arrangeGrob(p_lik, top = textGrob("A", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

# Label Delta K as 'b'
panel_b <- arrangeGrob(p_dk, top = textGrob("B", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

evanno_wbnp1 <- grid.arrange(panel_a, panel_b, 
                         ncol = 2, 
                         widths = c(1, 1))

ggsave("./output/plots/K_plots/w1.k.svg", 
       plot = evanno_wbnp1, 
       width = 27, 
       height = 11)

# WBNP2 ----
# K plots
w2.k <- evannoMethodStructure(data = struct_sum_w2, exportplot = F,
                           returnplot = T, returndata = F,
                           basesize = 25, linesize = 0.7,
                           height = 25, width = 25, units = "cm",
                           outputfilename = "w2_K",
                           exportpath = "./STRUCTURE_files/",
                           imgtype = "pdf")

p_lik <- w2.k$grobs[[1]]
p_dk <- w2.k$grobs[[4]]


# Label Likelihood as 'a'
panel_a <- arrangeGrob(p_lik, top = textGrob("A", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

# Label Delta K as 'b'
panel_b <- arrangeGrob(p_dk, top = textGrob("B", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

evanno_wbnp2 <- grid.arrange(panel_a, panel_b, 
                         ncol = 2, 
                         widths = c(1, 1))

ggsave("./output/plots/K_plots/w2.k.svg", 
       plot = evanno_wbnp2, 
       width = 27, 
       height = 11)

## ADMIX1 ----
# K plots
a1.k <- evannoMethodStructure(data = struct_sum_a1, exportplot = F,
                           returnplot = T, returndata = F,
                           basesize = 25, linesize = 0.7,
                           height = 25, width = 25, units = "cm",
                           outputfilename = "a1_K",
                           exportpath = "./STRUCTURE_files/",
                           imgtype = "pdf")

p_lik <- a1.k$grobs[[1]]
p_dk <- a1.k$grobs[[4]]


# Label Likelihood as 'a'
panel_a <- arrangeGrob(p_lik, top = textGrob("A", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

# Label Delta K as 'b'
panel_b <- arrangeGrob(p_dk, top = textGrob("B", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

evanno_admix1 <- grid.arrange(panel_a, panel_b, 
                         ncol = 2, 
                         widths = c(1, 1))

ggsave("./output/plots/K_plots/a1.k.svg", 
       plot = evanno_admix1, 
       width = 27, 
       height = 11)

## ADMIX2 ----
# K plots
a2.k <- evannoMethodStructure(data = struct_sum_a2, exportplot = F,
                           returnplot = T, returndata = F,
                           basesize = 25, linesize = 0.7,
                           height = 25, width = 25, units = "cm",
                           outputfilename = "a2_K",
                           exportpath = "./STRUCTURE_files/",
                           imgtype = "pdf")

p_lik <- a2.k$grobs[[1]]
p_dk <- a2.k$grobs[[4]]


# Label Likelihood as 'a'
panel_a <- arrangeGrob(p_lik, top = textGrob("A", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

# Label Delta K as 'b'
panel_b <- arrangeGrob(p_dk, top = textGrob("B", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

evanno_admix2 <- grid.arrange(panel_a, panel_b, 
                         ncol = 2, 
                         widths = c(1, 1))

ggsave("./output/plots/K_plots/a2.k.svg", 
       plot = evanno_admix2, 
       width = 27, 
       height = 11)
```

#### WBNP 1
```{r}
w1.data <- cleaned.data %>% 
  filter(Year.Collected <=2015) %>%
  rename(ID = Spreadsheet.ID) %>% 
  mutate(ID = as.character(ID)) %>% 
  filter(Herd %in% c("WBNPGR", "WBNPHC", "WBNPLB", "WBNPNY", "WBNPPAD", "WENT"))
# Align runs
slist_w1 <- alignK(slist_w1, "auto")

# Merge runs (average over 10 replicates)
w1.mk2 <- mergeQ(slist_w1[11:20])
w1.mk3 <- mergeQ(slist_w1[21:30])
w1.mk4 <- mergeQ(slist_w1[31:40])
w1.mk5 <- mergeQ(slist_w1[41:50])
w1.mk6 <- mergeQ(slist_w1[51:60])

w1.merged.list <- as.qlist(list(w1.mk2[[1]], w1.mk3[[1]], w1.mk4[[1]], 
                          w1.mk5[[1]], w1.mk6[[1]]))

ind_label_w1 <- readQ(files = sfiles_w1[1], indlabfrom = TRUE) %>% as.data.frame()
ind_label_w1 <- rownames_to_column(ind_label_w1, "ID")
locations_w1 <- left_join(ind_label_w1, w1.data, by = "ID")

loc_labels_w1 <- locations_w1$Herd %>% as.data.frame()

herd_order_w1 <- c("WBNPLB", "WBNPNY", "WBNPHC", "WBNPPAD", "WBNPGR","WENT")
struct_w1_cols <- c("#4477AA", "#EE6677",  "#66CCEE", "#AA3377", "#BBBBBB", "#EE7733")



plotQ(c(w1.mk2, w1.mk3, w1.mk4, w1.mk5, w1.mk6), 
      imgoutput = "join",
            basesize = 10, 
            grplab = loc_labels_w1, 
            subsetgrp = herd_order_w1, 
            clustercol = struct_w1_cols,
            ordergrp = TRUE,
      
      # Side "K" Labels
            splab = c("K = 2", "K = 3", "K = 4", "K = 5", "K = 6"),
            splabsize = 35,
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
            # Individual Label Settings
            showindlab = F, 
            indlabsize = 4,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 10, 
            grplabangle =45,
            grplabpos = 0.65,
            grplabheight = 1,      
            grplabspacer = 0.1,
            grplabcol = "black",
      
            linesize = 2,          
            pointsize = 15,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 3, 
            divtype = NULL,
      
            useindlab = TRUE, 
            returndata = TRUE,
      
      # Plot settings
            height = 7,             
            width = 50,
            panelspacer = 1,
            panelratio = c(1,1),
      
      
      # File settings
            imgtype = "tiff",
            outputfilename = "W1",
            exportpath = "./output/plots/structure_plots/wbnp/")
```

#### WBNP 2
```{r}
w2.data <- cleaned.data %>% 
  filter(Year.Collected >=2016) %>%
  rename(ID = Spreadsheet.ID) %>% 
  mutate(ID = as.character(ID)) %>% 
  filter(Herd %in% c("WBNPGR", "WBNPHC", "WBNPLB", "WBNPNY", "WBNPPAD", "WENT"))
# Align runs
slist_w2 <- alignK(slist_w2, "auto")

# Merge runs (average over 10 replicates)
w2.mk2 <- mergeQ(slist_w2[11:20])
w2.mk3 <- mergeQ(slist_w2[21:30])
w2.mk4 <- mergeQ(slist_w2[31:40])
w2.mk5 <- mergeQ(slist_w2[41:50])
w2.mk6 <- mergeQ(slist_w2[51:60])

w2.merged.list <- as.qlist(list(w2.mk2[[1]], w2.mk3[[1]], w2.mk4[[1]], 
                          w2.mk5[[1]], w2.mk6[[1]]))

ind_label_w2 <- readQ(files = sfiles_w2[1], indlabfrom = TRUE) %>% as.data.frame()
ind_label_w2 <- rownames_to_column(ind_label_w2, "ID")
locations_w2 <- left_join(ind_label_w2, w2.data, by = "ID")

loc_labels_w2 <- locations_w2$Herd %>% as.data.frame()

herd_order_w2 <- c("WBNPLB", "WBNPNY", "WBNPHC", "WBNPPAD", "WBNPGR","WENT")
struct_w2_cols <- c("#4477AA", "#EE6677",  "#66CCEE", "#AA3377", "#BBBBBB", "#EE7733")

plotQ(w2.merged.list, 
      imgoutput = "join",
            basesize = 10, 
            grplab = loc_labels_w2, 
            subsetgrp = herd_order_w2,
            clustercol = struct_w2_cols,
            ordergrp = TRUE,
      
      # Side "K" Labels
            splab = c("K = 2", "K = 3", "K = 4", "K = 5", "K = 6"),
            splabsize = 35,
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
            # Individual Label Settings
            showindlab = F, 
            indlabsize = 4,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 10, 
            grplabangle =45,
            grplabpos = 0.65,
            grplabheight = 1,      
            grplabspacer = 0.1,
            grplabcol = "black",
      
            linesize = 2,          
            pointsize = 15,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 3, 
            divtype = NULL,
      
            useindlab = TRUE, 
            returndata = TRUE,
      
      # Plot settings
            height = 7,             
            width = 50,
            panelspacer = 1,
            panelratio = c(1,1),
      
      
      # File settings
            imgtype = "tiff",
            outputfilename = "W2",
            exportpath = "./output/plots/structure_plots/wbnp")
```

#### ADMIX 1
```{r}
a1.data <- cleaned.data %>% 
  filter(Year.Collected <=2015) %>%
  rename(ID = Spreadsheet.ID) %>% 
  mutate(ID = as.character(ID)) %>% 
  filter(Herd %in% c("WBNPGR", "WBNPHC", "WBNPLB", "WBNPNY", "WBNPPAD", "WENT", "EINPW", "EINPP"))
# Align runs
slist_a1 <- alignK(slist_a1, "auto")

# Merge runs (average over 10 replicates)
a1.mk2 <- mergeQ(slist_a1[11:20])
a1.mk3 <- mergeQ(slist_a1[21:30])
a1.mk4 <- mergeQ(slist_a1[31:40])
a1.mk5 <- mergeQ(slist_a1[41:50])
a1.mk6 <- mergeQ(slist_a1[51:60])
a1.mk7 <- mergeQ(slist_a1[61:70])
a1.mk8 <- mergeQ(slist_a1[71:80])


a1.merged.list <- as.qlist(list(a1.mk2[[1]], a1.mk3[[1]], a1.mk4[[1]], 
                          a1.mk5[[1]], a1.mk6[[1]], a1.mk7[[1]], a1.mk8[[1]]))

ind_label_a1 <- readQ(files = sfiles_a1[1], indlabfrom = TRUE) %>% as.data.frame()
ind_label_a1 <- rownames_to_column(ind_label_a1, "ID")
locations_a1 <- left_join(ind_label_a1, a1.data, by = "ID")

loc_labels_a1 <- locations_a1$Herd %>% as.data.frame()

herd_order_a1 <- c("EINPP", "WBNPLB", "WBNPNY", "WBNPHC", "WBNPPAD", "WBNPGR","WENT", "EINPW")
struct_a1_cols <- c( "#228833","#CCBB44","#4477AA","#EE6677",  "#66CCEE", "#AA3377", "#BBBBBB", "#EE7733")

# Cluster ordering (colour matching/label switching)

# Set clusters for colour matching 
#mk2
Cluster1 <- a1.mk2$`2`$Cluster1 
Cluster2 <- a1.mk2$`2`$Cluster2 

a1.mk2$`2`$Cluster1 <- Cluster1
a1.mk2$`2`$Cluster2 <- Cluster2

#mk3
Cluster1 <- a1.mk3$`3`$Cluster1
Cluster3 <- a1.mk3$`3`$Cluster2 
Cluster2 <- a1.mk3$`3`$Cluster3 

a1.mk3$`3`$Cluster1 <- Cluster1
a1.mk3$`3`$Cluster2 <- Cluster2
a1.mk3$`3`$Cluster3 <- Cluster3

#mk4
Cluster1 <- a1.mk4$`4`$Cluster1
Cluster3 <- a1.mk4$`4`$Cluster2 
Cluster2 <- a1.mk4$`4`$Cluster3 
Cluster4 <- a1.mk4$`4`$Cluster4 

a1.mk4$`4`$Cluster1 <- Cluster1
a1.mk4$`4`$Cluster2 <- Cluster2
a1.mk4$`4`$Cluster3 <- Cluster3
a1.mk4$`4`$Cluster4 <- Cluster4

#mk5
Cluster1 <- a1.mk5$`5`$Cluster1
Cluster3 <- a1.mk5$`5`$Cluster2 
Cluster2 <- a1.mk5$`5`$Cluster3 
Cluster4 <- a1.mk5$`5`$Cluster4 
Cluster5 <- a1.mk5$`5`$Cluster5 

a1.mk5$`5`$Cluster1 <- Cluster1
a1.mk5$`5`$Cluster2 <- Cluster2
a1.mk5$`5`$Cluster3 <- Cluster3
a1.mk5$`5`$Cluster4 <- Cluster4
a1.mk5$`5`$Cluster5 <- Cluster5

#mk6
Cluster6 <- a1.mk6$`6`$Cluster1
Cluster3 <- a1.mk6$`6`$Cluster2 
Cluster2 <- a1.mk6$`6`$Cluster3 
Cluster4 <- a1.mk6$`6`$Cluster4 
Cluster5 <- a1.mk6$`6`$Cluster5 
Cluster1 <- a1.mk6$`6`$Cluster6 

a1.mk6$`6`$Cluster1 <- Cluster1
a1.mk6$`6`$Cluster2 <- Cluster2
a1.mk6$`6`$Cluster3 <- Cluster3
a1.mk6$`6`$Cluster4 <- Cluster4
a1.mk6$`6`$Cluster5 <- Cluster5
a1.mk6$`6`$Cluster6 <- Cluster6

#mk7
Cluster6 <- a1.mk7$`7`$Cluster1
Cluster3 <- a1.mk7$`7`$Cluster2 
Cluster2 <- a1.mk7$`7`$Cluster3 
Cluster4 <- a1.mk7$`7`$Cluster4 
Cluster5 <- a1.mk7$`7`$Cluster5 
Cluster1 <- a1.mk7$`7`$Cluster6 
Cluster7 <- a1.mk7$`7`$Cluster7 

a1.mk7$`7`$Cluster1 <- Cluster1
a1.mk7$`7`$Cluster2 <- Cluster2
a1.mk7$`7`$Cluster3 <- Cluster3
a1.mk7$`7`$Cluster4 <- Cluster4
a1.mk7$`7`$Cluster5 <- Cluster5
a1.mk7$`7`$Cluster6 <- Cluster6
a1.mk7$`7`$Cluster7 <- Cluster7

#mk8
Cluster6 <- a1.mk8$`8`$Cluster1
Cluster3 <- a1.mk8$`8`$Cluster2 
Cluster2 <- a1.mk8$`8`$Cluster3 
Cluster4 <- a1.mk8$`8`$Cluster4 
Cluster5 <- a1.mk8$`8`$Cluster5 
Cluster1 <- a1.mk8$`8`$Cluster6 
Cluster7 <- a1.mk8$`8`$Cluster7 
Cluster8 <- a1.mk8$`8`$Cluster8 

a1.mk8$`8`$Cluster1 <- Cluster1
a1.mk8$`8`$Cluster2 <- Cluster2
a1.mk8$`8`$Cluster3 <- Cluster3
a1.mk8$`8`$Cluster4 <- Cluster4
a1.mk8$`8`$Cluster5 <- Cluster5
a1.mk8$`8`$Cluster6 <- Cluster6
a1.mk8$`8`$Cluster7 <- Cluster7
a1.mk8$`8`$Cluster8 <- Cluster8


plotQ(c(a1.mk2, a1.mk3, a1.mk4, a1.mk5, a1.mk6, a1.mk7, a1.mk8), 
      imgoutput = "join",
            basesize = 10, 
            grplab = loc_labels_a1, 
            subsetgrp = herd_order_a1, 
            clustercol = struct_a1_cols,
            ordergrp = TRUE,
      
      # Side "K" Labels
            splab = c("K = 2", "K = 3", "K = 4", "K = 5", "K = 6", "K = 7", "K = 8"),
            splabsize = 35,
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
            # Individual Label Settings
            showindlab = F, 
            indlabsize = 4,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 10, 
            grplabangle =45,
            grplabpos = 0.65,
            grplabheight = 1,      
            grplabspacer = 0.1,
            grplabcol = "black",
      
            linesize = 2,          
            pointsize = 15,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 3, 
            divtype = NULL,
      
            useindlab = TRUE, 
            returndata = TRUE,
      
      # Plot settings
            height = 7,             
            width = 50,
            panelspacer = 1,
            panelratio = c(1,1),
      
      
      # File settings
            imgtype = "tiff",
            outputfilename = "A1",
            exportpath = "./output/plots/structure_plots/wbnp")
```


#### ADMIX 2
```{r}
a2.data <- cleaned.data %>% 
  filter(Year.Collected >=2016) %>%
  rename(ID = Spreadsheet.ID) %>% 
  mutate(ID = as.character(ID)) %>% 
  filter(Herd %in% c("WBNPGR", "WBNPHC", "WBNPLB", "WBNPNY", "WBNPPAD", "WENT", "EINPW", "EINPP"))
# Align runs
slist_a2 <- alignK(slist_a2, "auto")

# Merge runs (average over 10 replicates)
a2.mk2 <- mergeQ(slist_a2[11:20])
a2.mk3 <- mergeQ(slist_a2[21:30])
a2.mk4 <- mergeQ(slist_a2[31:40])
a2.mk5 <- mergeQ(slist_a2[41:50])
a2.mk6 <- mergeQ(slist_a2[51:60])
a2.mk7 <- mergeQ(slist_a2[61:70])
a2.mk8 <- mergeQ(slist_a2[71:80])


a2.merged.list <- as.qlist(list(a2.mk2[[1]], a2.mk3[[1]], a2.mk4[[1]], 
                          a2.mk5[[1]], a2.mk6[[1]], a2.mk7[[1]], a2.mk8[[1]]))

ind_label_a2 <- readQ(files = sfiles_a2[1], indlabfrom = TRUE) %>% as.data.frame()
ind_label_a2 <- rownames_to_column(ind_label_a2, "ID")
locations_a2 <- left_join(ind_label_a2, a2.data, by = "ID")

loc_labels_a2 <- locations_a2$Herd %>% as.data.frame()

herd_order_a2 <- c("EINPP", "WBNPLB", "WBNPNY", "WBNPHC", "WBNPPAD", "WBNPGR","WENT", "EINPW")
struct_a2_cols <- c( "#228833","#CCBB44","#4477AA","#EE6677",  "#66CCEE", "#AA3377", "#BBBBBB", "#EE7733")

# Set clusters for colour matching 
#mk2
Cluster2 <- a2.mk2$`2`$Cluster1 
Cluster1 <- a2.mk2$`2`$Cluster2 

a2.mk2$`2`$Cluster1 <- Cluster1
a2.mk2$`2`$Cluster2 <- Cluster2

#mk3
Cluster3 <- a2.mk3$`3`$Cluster1
Cluster1 <- a2.mk3$`3`$Cluster2 
Cluster2 <- a2.mk3$`3`$Cluster3 

a2.mk3$`3`$Cluster1 <- Cluster1
a2.mk3$`3`$Cluster2 <- Cluster2
a2.mk3$`3`$Cluster3 <- Cluster3

#mk4
Cluster3 <- a2.mk4$`4`$Cluster1
Cluster1 <- a2.mk4$`4`$Cluster2 
Cluster2 <- a2.mk4$`4`$Cluster3 
Cluster4 <- a2.mk4$`4`$Cluster4 

a2.mk4$`4`$Cluster1 <- Cluster1
a2.mk4$`4`$Cluster2 <- Cluster2
a2.mk4$`4`$Cluster3 <- Cluster3
a2.mk4$`4`$Cluster4 <- Cluster4

#mk5
Cluster3 <- a2.mk5$`5`$Cluster1
Cluster1 <- a2.mk5$`5`$Cluster2 
Cluster2 <- a2.mk5$`5`$Cluster3 
Cluster4 <- a2.mk5$`5`$Cluster4 
Cluster5 <- a2.mk5$`5`$Cluster5 

a2.mk5$`5`$Cluster1 <- Cluster1
a2.mk5$`5`$Cluster2 <- Cluster2
a2.mk5$`5`$Cluster3 <- Cluster3
a2.mk5$`5`$Cluster4 <- Cluster4
a2.mk5$`5`$Cluster5 <- Cluster5

#mk6
Cluster3 <- a2.mk6$`6`$Cluster1
Cluster1 <- a2.mk6$`6`$Cluster2 
Cluster2 <- a2.mk6$`6`$Cluster3 
Cluster4 <- a2.mk6$`6`$Cluster4 
Cluster5 <- a2.mk6$`6`$Cluster5 
Cluster6 <- a2.mk6$`6`$Cluster6 

a2.mk6$`6`$Cluster1 <- Cluster1
a2.mk6$`6`$Cluster2 <- Cluster2
a2.mk6$`6`$Cluster3 <- Cluster3
a2.mk6$`6`$Cluster4 <- Cluster4
a2.mk6$`6`$Cluster5 <- Cluster5
a2.mk6$`6`$Cluster6 <- Cluster6

#mk7
Cluster3 <- a2.mk7$`7`$Cluster1
Cluster1 <- a2.mk7$`7`$Cluster2 
Cluster2 <- a2.mk7$`7`$Cluster3 
Cluster4 <- a2.mk7$`7`$Cluster4 
Cluster5 <- a2.mk7$`7`$Cluster5 
Cluster6 <- a2.mk7$`7`$Cluster6 
Cluster7 <- a2.mk7$`7`$Cluster7 

a2.mk7$`7`$Cluster1 <- Cluster1
a2.mk7$`7`$Cluster2 <- Cluster2
a2.mk7$`7`$Cluster3 <- Cluster3
a2.mk7$`7`$Cluster4 <- Cluster4
a2.mk7$`7`$Cluster5 <- Cluster5
a2.mk7$`7`$Cluster6 <- Cluster6
a2.mk7$`7`$Cluster7 <- Cluster7

#mk8
Cluster3 <- a2.mk8$`8`$Cluster1
Cluster1 <- a2.mk8$`8`$Cluster2 
Cluster2 <- a2.mk8$`8`$Cluster3 
Cluster4 <- a2.mk8$`8`$Cluster4 
Cluster5 <- a2.mk8$`8`$Cluster5 
Cluster6 <- a2.mk8$`8`$Cluster6 
Cluster7 <- a2.mk8$`8`$Cluster7 
Cluster8 <- a2.mk8$`8`$Cluster8 

a2.mk8$`8`$Cluster1 <- Cluster1
a2.mk8$`8`$Cluster2 <- Cluster2
a2.mk8$`8`$Cluster3 <- Cluster3
a2.mk8$`8`$Cluster4 <- Cluster4
a2.mk8$`8`$Cluster5 <- Cluster5
a2.mk8$`8`$Cluster6 <- Cluster6
a2.mk8$`8`$Cluster7 <- Cluster7
a2.mk8$`8`$Cluster8 <- Cluster8


plotQ(c(a2.mk2, a2.mk3, a2.mk4, a2.mk5, a2.mk6, a2.mk7, a2.mk8), 
      imgoutput = "join",
            basesize = 10, 
            grplab = loc_labels_a2, 
            subsetgrp = herd_order_a2, 
            clustercol = struct_a2_cols,
            ordergrp = TRUE,
      
      # Side "K" Labels
            splab = c("K = 2", "K = 3", "K = 4", "K = 5", "K = 6", "K = 7", "K = 8"),
            splabsize = 35,
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
            # Individual Label Settings
            showindlab = F, 
            indlabsize = 4,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 10, 
            grplabangle =45,
            grplabpos = 0.65,
            grplabheight = 1,      
            grplabspacer = 0.1,
            grplabcol = "black",
      
            linesize = 2,          
            pointsize = 15,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 3, 
            divtype = NULL,
      
            useindlab = TRUE, 
            returndata = TRUE,
      
      # Plot settings
            height = 7,             
            width = 50,
            panelspacer = 1,
            panelratio = c(1,1),
      
      
      # File settings
            imgtype = "tiff",
            outputfilename = "A2",
            exportpath = "./output/plots/structure_plots/wbnp")
```

# DAPC

#### WBNP samples temp1
```{r echo=TRUE}
## Unsupervised K means clustering
# Identify K Clusters 
grp1 <- find.clusters(w1, max.n.clust = 8, n.pca = (nInd(w1) -1))

# Print BIC plot
svg(file = "./output/plots/DAPC/w1_BIC.svg", width = 7, height = 5)
plot(grp1$Kstat, type = "o", xlab = "Number of clusters (K)", 
     ylab = "BIC", main = "Value of BIC \n versus number of clusters \n WBNP: 1986 - 2015" ,
     col = "blue", pch = 1)
dev.off()

# BIC minimum value is at K =1
table.value(table(pop(w1), grp1$grp), col.lab=paste("inf", 1:6),
row.lab=paste("ori", 1:6))

## Supervised DAPC
# TO prevent overfitting of the model, find optimal PCs to retain (Max = n-1 individuals)
set.seed(999)
pramx <- xvalDapc(tab(w1, NA.method = "mean"), pop(w1))

# Determine optimal number of PCs
numpc1 <- as.numeric(pramx$`Number of PCs Achieving Highest Mean Success`)

# DAPC with known parameters
dapc1 <- dapc(w1, pop(w1), n.pca = numpc1, n.da = (length(pop(w1))-1))
dapc1
```

#### WBNP Samples temp 2
```{r echo=TRUE}
## Unsupervised K means clustering
# Identify K Clusters 
grp2 <- find.clusters(w2, max.n.clust = 8, n.pca = (nInd(w2) -1))

# Print BIC plot
svg(file = "./output/plots/DAPC/w2_BIC.svg", width = 7, height = 5)
plot(grp2$Kstat, type = "o", xlab = "Number of clusters (K)", 
     ylab = "BIC", main = "Value of BIC \n versus number of clusters \n WBNP: 2016 - 2026" ,
     col = "blue", pch = 1)
dev.off()

# BIC minimum value is at K =1
table.value(table(pop(w2), grp2$grp), col.lab=paste("inf", 1:6),
row.lab=paste("ori", 1:6))

## Supervised DAPC
# TO prevent overfitting of the model, find optimal PCs to retain (Max = n-1 individuals)
set.seed(999)
pramx2 <- xvalDapc(tab(w2, NA.method = "mean"), pop(w2),
                   xval.plot = TRUE)

# Determine optimal number of PCs
numpc2 <- as.numeric(pramx2$`Number of PCs Achieving Highest Mean Success`)

# DAPC with known parameters
dapc2 <- dapc(w2, pop(w2), n.pca = numpc2, n.da = (length(pop(w2))-1))
dapc2
```

#### Plots BIC plots
```{r}
####### Plot ----
# Plot
svg(file = "./output/plots/DAPC/wbnp_BIC_merged.svg", width = 10, height = 5) 

# Set up  grid 1 row, 2 columns
par(mfrow = c(1, 2)) 

# Temp1
plot(grp1$Kstat, type = "o", xlab = "Number of clusters (K)", 
     ylab = "BIC", main = "Value of BIC \n versus number of clusters \n WBNP: 1986 - 2015" ,
     col = "blue", pch = 1)

# Temp2
plot(grp2$Kstat, type = "o", xlab = "Number of clusters (K)", 
     ylab = "BIC", main = "Value of BIC \n versus number of clusters \n WBNP: 2016 - 2026" ,
     col = "blue", pch = 1)


dev.off()
```


```{r}
####### Plot ----
dapc_palette <- c("#cecece", 
                  "#a559aa", 
                  "#59a89c", 
                  "#f0c571", 
                  "#e02b35", 
                  "#082a54"
                  )

# Plot
svg(file = "./output/plots/DAPC/wbnp_DAPC_stacked.svg", width = 20, height = 10) 

# Set up  grid 1 row, 2 columns
par(mfrow = c(1, 2)) 

# Plot 1
scatter(dapc1,posi.da="bottomright", bg="white",
pch=17:22,cstar=0, cex = 2, col=dapc_palette,scree.pca=TRUE,
posi.pca = "topright", ratio.da = 0.20, ratio.pca = 0.13, csub = 2,
sub = paste0("DA1: ", round(dapc1$eig[1], 1), "% | DA2: ", round(dapc1$eig[2], 1), "%"))
mtext("A", side = 3, adj = -7.8, line = 2, cex = 2, font = 2)

# Plot 2
scatter(dapc2,posi.da="bottomright", bg="white",
pch=17:22,cstar=0, cex = 2,col=dapc_palette,scree.pca=TRUE,
posi.pca = "topright", ratio.da = 0.20, ratio.pca = 0.13, csub = 2,
sub = paste0("DA1: ", round(dapc2$eig[1], 1), "% | DA2: ", round(dapc2$eig[2], 1), "%"))
mtext("B", side = 3, adj = -7.8, line = 2, cex = 2, font = 2)

dev.off()
```







#### SINGLE ADMIXTURE PLOTS
```{r}
herd_order <- c("EINPP","WBNPLB", "WBNPNY", "WBNPHC", "WBNPPAD", "WBNPGR","WENT", "EINPW")
master_cols <- c( "#228833","#CCBB44","#4477AA","#EE6677",  "#66CCEE", "#AA3377", "#BBBBBB", "#EE7733")


# Flip clusters so they plot correctly
#a1.mk2$`2` <- a1.mk2$`2`[, c("Cluster2", "Cluster1")]
#colnames(a1.mk2$`2`) <- c("Cluster1", "Cluster2")

# STRUCTURE A1 with EINP outgroups
admixture1 <- plotQ(a1.mk2,
            #basesize = 50, 
            grplab = loc_labels_a1, 
            subsetgrp = herd_order, 
            clustercol = master_cols,
            ordergrp = TRUE,

      # Side "K" Labels
            splab = "K=2",
            splabsize = 20,
            splabface = "bold",
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
      # Individual Label Settings
            showindlab = F, 
            indlabsize = 0,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 6.2, 
            grplabangle = 45,
            grplabpos = 0.85,
            grplabheight = 0.6,      
            grplabspacer = 0,
            grplabcol = "black",
      
            linesize = 1,          
            pointsize = 8,
            linepos = 0.95,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 1, 
            divtype = NULL,
            useindlab = TRUE, 
            returndata = TRUE,
            #barbordersize = 0.01,
            #barbordercolour = "lightgrey",
      
      # Add title
           showtitle = TRUE,
           titlelab = NULL,
           titlehjust = 0.5,
           titlevjust = 0.5,
           titlesize = 15,
           titlecol = "black",
           titleface = "bold",
      
      # Plot settings
            height = 10,             
            width = 35,
            panelspacer = 0.2,
            panelratio = c(1,2),
      
      # File settings
            imgtype = "tiff",
            returnplot = TRUE,
            outputfilename = "plota1.k2",
            exportpath = "./output/plots/structure_plots/wbnp/")

# Second Temporal Group

# Flip clusters so they plot correctly
#a2.mk2$`2` <- a2.mk2$`2`[, c("Cluster2", "Cluster1")]
#colnames(a2.mk2$`2`) <- c("Cluster1", "Cluster2")

# STRUCTURE A1 with EINP outgroups
admixture2 <- plotQ(a2.mk2,
            #basesize = 20, 
            grplab = loc_labels_a2, 
            subsetgrp = herd_order, 
            clustercol = master_cols,
            ordergrp = TRUE,

      # Side "K" Labels
            splab = "K=2",
            splabsize = 20,
            splabface = "bold",
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
      # Individual Label Settings
            showindlab = F, 
            indlabsize = 0,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 6.2, 
            grplabangle = 45,
            grplabpos = 0.85,
            grplabheight = 0.6,      
            grplabspacer = 0,
            grplabcol = "black",
      
            linesize = 1,          
            pointsize = 8,
            linepos = 0.95,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 1, 
            divtype = NULL,
            useindlab = TRUE, 
            returndata = TRUE,
            #barbordersize = 0.01,
            #barbordercolour = "lightgrey",

      # Add title
           showtitle = TRUE,
           titlelab = NULL,
           titlehjust = 0.5,
           titlevjust = 0.5,
           titlesize = 15,
           titlecol = "black",
      titleface = "bold",
      
      # Plot settings
            height = 10,             
            width = 35,
            panelspacer = 0.2,
            panelratio = c(1,2),
      
      # File settings
            imgtype = "tiff",
            returnplot = TRUE,
            outputfilename = "plota2.k2",
            exportpath = "./output/plots/structure_plots/wbnp/")
```

# Plains Admixture
#### Assess Q-values (membership probabilities of assigning to clusters)
```{r}
a1.order <- data.frame(admixture1$data$grplab[[1]]) %>% 
  rownames_to_column("Ind")

a1.k2 <- data.frame(a1.mk2) %>% 
  rownames_to_column("Ind") %>% 
  left_join(a1.order, by = "Ind") %>% 
  rename(Herd = ".",
        Wood = X2.Cluster1,
         Plains = X2.Cluster2) %>% 
  select(-Ind) %>% 
  relocate(Herd, .before = Plains)
  
a1.Q.sum <- a1.k2 %>% 
  group_by(Herd) %>% 
  summarise(
    QWood = paste0(round(mean(Wood), 3), "±", round(sd(Wood), 3)),
    QPlains = paste0(round(mean(Plains), 3), "±", round(sd(Plains), 3))
  )

# Summary per herd of Q assignments
a1.Q.sum

# Just WBNP
a1.wbnp.Q.sum <-a1.Q.sum[3:8, ] 
mean(a1.wbnp.Q.sum$Plains_mean)

write.csv(a1.wbnp.Q.sum, "./output/tables/structure/a1_PB_assignment.csv")

#-------------------


a2.order <- data.frame(admixture2$data$grplab[[1]]) %>% 
  rownames_to_column("Ind")

a2.k2 <- data.frame(a2.mk2) %>% 
  rownames_to_column("Ind") %>% 
  left_join(a2.order, by = "Ind") %>% 
  rename(Herd = ".",
         Wood = X2.Cluster1, # Swapped from previous
         Plains = X2.Cluster2) %>% 
  select(-Ind) %>% 
  relocate(Herd, .before = Wood)
  
a2.Q.sum <- a2.k2 %>% 
  group_by(Herd) %>% 
  summarise(
    QWood = paste0(round(mean(Wood), 3), "±", round(sd(Wood), 3)),
    QPlains = paste0(round(mean(Plains), 3), "±", round(sd(Plains), 3))
  )

# Summary per herd of Q assignments
a2.Q.sum

a2.wbnp.Q.sum <-a2.Q.sum[3:8, ] 
mean(a2.wbnp.Q.sum$Plains_mean)


write.csv(a2.wbnp.Q.sum, "./output/tables/structure/a2_PB_assignment.csv")


```


#### JUST WBNP Plains (but grouped as all herds)
```{r}
# Identify the names of the WBNP herds based on your subsetting (rows 3 to 8)
wbnp_herds <- a1.Q.sum$Herd[3:8]

# Filter the original individual data for these herds and calculate the overall stats
a1.wbnp.overall_stats <- a1.k2 %>%
  filter(Herd %in% wbnp_herds) %>%
  summarise(
    Overall_Mean_Plains = mean(Plains, na.rm = TRUE),
    Overall_SD_Plains = sd(Plains, na.rm = TRUE)
  )

# View the result
a1.wbnp.overall_stats


# Filter the original individual data for these herds and calculate the overall stats
a2.wbnp.overall_stats <- a2.k2 %>%
  filter(Herd %in% wbnp_herds) %>%
  summarise(
    Overall_Mean_Plains = mean(Plains, na.rm = TRUE),
    Overall_SD_Plains = sd(Plains, na.rm = TRUE)
  )

# View the result
a2.wbnp.overall_stats
```



######### DOWNSAMPLE WBNP T2
```{r}
#### Data Subset and Genind Creation -------------------------------------------
# Set seed
set.seed(123)
# Assign population sizes to downsample for PCA and Structure analyses
downsample <- tibble(
  Herd = c("WBNPGR", "WBNPHC", "WBNPLB", "WBNPNY", "WBNPPAD", "WENT"),
  n = c(10, 23, 11, 18, 25, 5)
)

downsample_data <- cleaned.data %>% 
  filter(Herd %in% c("WBNPGR", "WBNPHC", "WBNPLB", "WBNPNY", "WBNPPAD", "WENT") &
           Year.Collected >= 2016) %>% 
  left_join(downsample, by = "Herd") %>% 
  relocate(n, .before = Herd) %>% 
  group_by(Herd) %>% 
  reframe(
    slice_sample(
      pick(everything()),
      n = min(nrow(cur_data()), unique(n))
    )
  )
# Downsample genind creation
# Make genind object
downsample.genind <- df2genind(X=downsample_data[,c(16:ncol(downsample_data))],
                              sep = "/",
                              ind.names = downsample_data$Spreadsheet.ID,
                              loc.names = NULL,
                              pop = downsample_data$Herd,
                              NA.char = "NA",
                              ploidy = 2,
                              type = "codom",
                              strata = NULL,
                              hierarchy = NULL)

# Attach collection year metadata
downsample.genind@other$year <- downsample_data$Year.Collected

table(pop(downsample.genind))


#### PCA -----------------------------------------------------------------------
x.subw2 <- tab(downsample.genind, freq = TRUE, NA.method = "mean")
pca.subw2 <- dudi.pca(x.subw2, center = T, scale = F, scannf = F, nf = 2)

# Axes % Variance
eig.perc.subw2 <- 100*pca.subw2$eig/sum(pca.subw2$eig)

pca.subw2$li[ ,2] <- pca.subw2$li[ ,2] * -1


# Plot PCA object
pcasubw2 <- fviz_pca_ind(pca.subw2,
             axes = c(1, 2),
             geom = "blank",
             habillage = downsample.genind@pop,
             repel = TRUE,
             palette = w2_cols,
             alpha.ind = 1) + 
  stat_ellipse(aes(color = downsample.genind@pop), type = "norm", level = 0.68) + 
  geom_point(aes(colour = downsample.genind@pop),
             shape = 19,
             alpha = 0.75,
             size = 1.2) + 
  coord_fixed(ratio = 1, xlim = c(-2,2), ylim = c(-2, 2)) +
  
theme(
    plot.title = element_blank(),
    legend.title = element_text(face = "bold", size = 15),
    legend.text = element_text(size = 15),
    axis.title = element_text(face = "bold", size = 20),
    text = element_text(size = 20),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),  ) + 


  labs(title = NULL,
       x = paste0("PC1 (", round(eig.perc.subw2[1], 2), "%)"),
       y = paste0("PC2 (", round(eig.perc.subw2[2], 2), "%)"),
       colour = "Herd") + 
    guides(colour = guide_legend(override.aes = list(size = 4, shape = 19)))

pcasubw2

ggsave("./output/plots/PCAs/wbnp/downsampled_temp2_wbnp.svg",
       plot = pcasubw2, 
       device = "svg", 
       width = 12, 
       height = 12, 
       units = "in")


#### Discriminant Analysis of Principal Components -----------------------------
## Unsupervised K means clustering
# Identify K Clusters 
grpsub2 <- find.clusters(downsample.genind, max.n.clust = 8, n.pca = (nInd(downsample.genind) -1))

# Print BIC plot
svg(file = "./output/plots/DAPC/subw2_BIC.svg", width = 7, height = 5)
plot(grpsub2$Kstat, type = "o", xlab = "Number of clusters (K)", 
     ylab = "BIC", main = "Value of BIC \n versus number of clusters \n Subsampled WBNP: 2016 - 2026" ,
     col = "blue", pch = 1)
dev.off()

# BIC minimum value is at K = 2
table.value(table(pop(downsample.genind), grpsub2$grp), col.lab=paste("inf", 1:6),
row.lab=paste("ori", 1:6))

## Supervised DAPC
# TO prevent overfitting of the model, find optimal PCs to retain (Max = n-1 individuals)
set.seed(999)
pramxsub2 <- xvalDapc(tab(downsample.genind, NA.method = "mean"), pop(downsample.genind),
                   xval.plot = TRUE)

# Determine optimal number of PCs
numpcsub2 <- as.numeric(pramxsub2$`Number of PCs Achieving Highest Mean Success`)

# DAPC with known parameters
dapcsub2 <- dapc(downsample.genind, pop(downsample.genind), n.pca = numpcsub2, n.da = (length(pop(w2))-1))
dapcsub2

# Plot subw2
svg(file = "./output/plots/DAPC/subw2_DAPC.svg", width = 7, height = 5)
scatter(dapcsub2,posi.da="bottomright", bg="white",
pch=17:22,cstar=0, cex = 2,col=dapc_palette,scree.pca=TRUE,
posi.pca = "topright", ratio.da = 0.20, ratio.pca = 0.13, csub = 2,
sub = paste0("DA1: ", round(dapcsub2$eig[1], 1), "% | DA2: ", round(dapcsub2$eig[2], 1), "%"))
dev.off()

#### STRUCTURE analyses --------------------------------------------------------
# Create Structure input files
# subW2 structure (93 individuals, 6 pops)
sub_wbnp2 <- downsample.genind
pop(sub_wbnp2) <- as.numeric(as.factor(pop(sub_wbnp2)))

genind_to_structure(sub_wbnp2,
                      output = "./STRUCTURE/wbnp/subw2.structure.txt")

## Import files
# List files
# Full datasets
sfiles_subw2 <- list.files("./STRUCTURE/wbnp/downsample/results/", full.names = TRUE)

# Qlist (pophelper)
slist_subw2 <- readQ(files = sfiles_subw2, filetype = "structure")

# Summarise runs
struct_sum_subw2 <- summariseQ(tabulateQ(slist_subw2))

### Evanno Plots
# K plots
subw2.k <- evannoMethodStructure(data = struct_sum_subw2, exportplot = F,
                           returnplot = T, returndata = F,
                           basesize = 25, linesize = 0.7,
                           height = 25, width = 25, units = "cm",
                           outputfilename = "subw2_K",
                           exportpath = "./output/plots/K_plots/",
                           imgtype = "pdf")

p_lik <- subw2.k$grobs[[1]]
p_dk <- subw2.k$grobs[[4]]


# Label Likelihood as 'a'
panel_a <- arrangeGrob(p_lik, top = textGrob("A", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

# Label Delta K as 'b'
panel_b <- arrangeGrob(p_dk, top = textGrob("B", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

evanno_wbnpsub2 <- grid.arrange(panel_a, panel_b, 
                         ncol = 2, 
                         widths = c(1, 1))

ggsave("./output/plots/K_plots/subw2.k.svg", 
       plot = evanno_wbnpsub2, 
       width = 27, 
       height = 11)
```

```{r}
## Structure Plots
subw2.data <-  downsample_data %>% 
    rename(ID = Spreadsheet.ID) %>% 
    mutate(ID = as.character(ID))
# Align runs
slist_subw2 <- alignK(slist_subw2, "auto")

# Merge runs (average over 10 replicates)
subw2.mk2 <- mergeQ(slist_subw2[11:20])
subw2.mk3 <- mergeQ(slist_subw2[21:30])
subw2.mk4 <- mergeQ(slist_subw2[31:40])
subw2.mk5 <- mergeQ(slist_subw2[41:50])
subw2.mk6 <- mergeQ(slist_subw2[51:60])

subw2.merged.list <- as.qlist(list(subw2.mk2[[1]], subw2.mk3[[1]], subw2.mk4[[1]], 
                          subw2.mk5[[1]], subw2.mk6[[1]]))

ind_label_subw2 <- readQ(files = sfiles_subw2[1], indlabfrom = TRUE) %>% as.data.frame()
ind_label_subw2 <- rownames_to_column(ind_label_subw2, "ID")
locations_subw2 <- left_join(ind_label_subw2, subw2.data, by = "ID")

loc_labels_subw2 <- locations_subw2$Herd %>% as.data.frame()

herd_order_subw2 <- c("WBNPLB", "WBNPNY", "WBNPHC", "WBNPPAD", "WBNPGR","WENT")
struct_subw2_cols <- struct_a1_cols <- c("#4477AA","#EE6677",  "#66CCEE", "#AA3377", "#BBBBBB", "#EE7733")


# Set clusters for colour matching 
#mk2
#subw2.mk2$`2` <- subw2.mk2$`2`[, c("Cluster2", "Cluster1")]
#colnames(subw2.mk2$`2`) <- c("Cluster1", "Cluster2")

Cluster2 <- subw2.mk2$`2`$Cluster1 
Cluster1 <- subw2.mk2$`2`$Cluster2 

subw2.mk2$`2`$Cluster1 <- Cluster1
subw2.mk2$`2`$Cluster2 <- Cluster2

#mk3
#subw2.mk3$`3` <- subw2.mk3$`3`[, c("Cluster2", "Cluster1", "Cluster3")]
#colnames(subw2.mk3$`3`) <- c("Cluster1", "Cluster2", "Cluster3")

Cluster2 <- subw2.mk3$`3`$Cluster1
Cluster1 <- subw2.mk3$`3`$Cluster2 
Cluster3 <- subw2.mk3$`3`$Cluster3 

subw2.mk3$`3`$Cluster1 <- Cluster1
subw2.mk3$`3`$Cluster2 <- Cluster2
subw2.mk3$`3`$Cluster3 <- Cluster3

#mk4
Cluster2 <- subw2.mk4$`4`$Cluster1
Cluster1 <- subw2.mk4$`4`$Cluster2 
Cluster3 <- subw2.mk4$`4`$Cluster3 
Cluster4 <- subw2.mk4$`4`$Cluster4 

subw2.mk4$`4`$Cluster1 <- Cluster1
subw2.mk4$`4`$Cluster2 <- Cluster2
subw2.mk4$`4`$Cluster3 <- Cluster3
subw2.mk4$`4`$Cluster4 <- Cluster4

#mk5
Cluster1 <- subw2.mk5$`5`$Cluster1
Cluster2 <- subw2.mk5$`5`$Cluster2 
Cluster3 <- subw2.mk5$`5`$Cluster3 
Cluster4 <- subw2.mk5$`5`$Cluster4 
Cluster5 <- subw2.mk5$`5`$Cluster5 

subw2.mk5$`5`$Cluster1 <- Cluster1
subw2.mk5$`5`$Cluster2 <- Cluster2
subw2.mk5$`5`$Cluster3 <- Cluster3
subw2.mk5$`5`$Cluster4 <- Cluster4
subw2.mk5$`5`$Cluster5 <- Cluster5

#mk6
Cluster1 <- subw2.mk6$`6`$Cluster1
Cluster2 <- subw2.mk6$`6`$Cluster2 
Cluster3 <- subw2.mk6$`6`$Cluster3 
Cluster4 <- subw2.mk6$`6`$Cluster4 
Cluster5 <- subw2.mk6$`6`$Cluster5 
Cluster6 <- subw2.mk6$`6`$Cluster6 

subw2.mk6$`6`$Cluster1 <- Cluster1
subw2.mk6$`6`$Cluster2 <- Cluster2
subw2.mk6$`6`$Cluster3 <- Cluster3
subw2.mk6$`6`$Cluster4 <- Cluster4
subw2.mk6$`6`$Cluster5 <- Cluster5
subw2.mk6$`6`$Cluster6 <- Cluster6

plotQ(c(subw2.mk2, subw2.mk3, subw2.mk4, subw2.mk5, subw2.mk6), 
      imgoutput = "join",
            basesize = 10, 
            grplab = loc_labels_subw2, 
            subsetgrp = herd_order_subw2,
            clustercol = struct_subw2_cols,
            ordergrp = TRUE,
      
      # Side "K" Labels
            splab = c("K = 2", "K = 3", "K = 4", "K = 5", "K = 6"),
            splabsize = 35,
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
            # Individual Label Settings
            showindlab = F, 
            indlabsize = 4,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 10, 
            grplabangle =45,
            grplabpos = 0.65,
            grplabheight = 1,      
            grplabspacer = 0.1,
            grplabcol = "black",
      
            linesize = 2,          
            pointsize = 15,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 3, 
            divtype = NULL,
      
            useindlab = TRUE, 
            returndata = TRUE,
      
      # Plot settings
            height = 7,             
            width = 50,
            panelspacer = 1,
            panelratio = c(1,1),
      
      
      # File settings
            imgtype = "tiff",
            outputfilename = "subW2",
            exportpath = "./output/plots/structure_plots/wbnp")
```
## WBNP 2 clusters

```{r}
# Temp1
plotQ(w1.mk2, 
            basesize = 10, 
            grplab = loc_labels_w1, 
            subsetgrp = herd_order_w1,
            clustercol = struct_w1_cols,
            ordergrp = TRUE,

      # Side "K" Labels
            splab = "K=2",
            splabsize = 20,
            splabface = "bold",
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
      # Individual Label Settings
            showindlab = F, 
            indlabsize = 0,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 6.2, 
            grplabangle = 45,
            grplabpos = 0.85,
            grplabheight = 0.6,      
            grplabspacer = 0,
            grplabcol = "black",
      
            linesize = 1,          
            pointsize = 8,
            linepos = 0.95,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 1, 
            divtype = NULL,
            useindlab = TRUE, 
            returndata = TRUE,
            #barbordersize = 0.01,
            #barbordercolour = "lightgrey",
      
      # Add title
           showtitle = TRUE,
           titlelab = NULL,
           titlehjust = 0.5,
           titlevjust = 0.5,
           titlesize = 15,
           titlecol = "black",
           titleface = "bold",
      
      # Plot settings
            height = 10,             
            width = 35,
            panelspacer = 0.2,
            panelratio = c(1,2),
      
      
      # File settings
            imgtype = "tiff",
            outputfilename = "w1.k2",
            exportpath = "./output/plots/structure_plots/wbnp")


plotw2 <- plotQ(w2.mk2, 
            basesize = 10, 
            grplab = loc_labels_w2, 
            subsetgrp = herd_order_w2,
            clustercol = struct_w2_cols,
            ordergrp = TRUE,
      

      # Side "K" Labels
            splab = "K=2",
            splabsize = 20,
            splabface = "bold",
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
      # Individual Label Settings
            showindlab = F, 
            indlabsize = 0,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 6.2, 
            grplabangle = 45,
            grplabpos = 0.85,
            grplabheight = 0.6,      
            grplabspacer = 0,
            grplabcol = "black",
      
            linesize = 1,          
            pointsize = 8,
            linepos = 0.95,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 1, 
            divtype = NULL,
            useindlab = TRUE, 
            returndata = TRUE,
            #barbordersize = 0.01,
            #barbordercolour = "lightgrey",
      
      # Add title
           showtitle = TRUE,
           titlelab = NULL,
           titlehjust = 0.5,
           titlevjust = 0.5,
           titlesize = 15,
           titlecol = "black",
           titleface = "bold",
      
      # Plot settings
            height = 10,             
            width = 35,
            panelspacer = 0.2,
            panelratio = c(1,2),
      returnplot = TRUE,
      
      
      # File settings
            imgtype = "tiff",
            outputfilename = "w2.k2",
            exportpath = "./output/plots/structure_plots/wbnp")


## Downsampled W2

# Flip clusters so they plot correctly
#subw2.mk2$`2` <- subw2.mk2$`2`[, c("Cluster2", "Cluster1")]
#colnames(subw2.mk2$`2`) <- c("Cluster1", "Cluster2")

plotQ(subw2.mk2, 
            basesize = 10, 
            grplab = loc_labels_subw2, 
            subsetgrp = herd_order_subw2,
            clustercol = struct_subw2_cols,
            ordergrp = TRUE,
      

      # Side "K" Labels
            splab = "K=2",
            splabsize = 20,
            splabface = "bold",
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
      # Individual Label Settings
            showindlab = F, 
            indlabsize = 0,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 6.2, 
            grplabangle = 45,
            grplabpos = 0.85,
            grplabheight = 0.6,      
            grplabspacer = 0,
            grplabcol = "black",
      
            linesize = 1,          
            pointsize = 8,
            linepos = 0.95,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 1, 
            divtype = NULL,
            useindlab = TRUE, 
            returndata = TRUE,
            #barbordersize = 0.01,
            #barbordercolour = "lightgrey",
      
      # Add title
           showtitle = TRUE,
           titlelab = NULL,
           titlehjust = 0.5,
           titlevjust = 0.5,
           titlesize = 15,
           titlecol = "black",
           titleface = "bold",
      
      # Plot settings
            height = 10,             
            width = 35,
            panelspacer = 0.2,
            panelratio = c(1,2),
      
      # File settings
            imgtype = "tiff",
            outputfilename = "subW2.k2",
            exportpath = "./output/plots/structure_plots/wbnp")
```

# WBNP Split
Based on clustering analyses, we see two putative genetic clusters. What herds assign to what cluster:
```{r}
w2.order <- data.frame(plotw2$data$grplab[[1]]) %>% 
  rownames_to_column("Ind")

w2.k2 <- data.frame(w2.mk2) %>% 
  rownames_to_column("Ind") %>% 
  left_join(w2.order, by = "Ind") %>% 
  rename(Herd = ".",
        North = X2.Cluster2,
         Central = X2.Cluster1) %>% 
  select(-Ind) %>% 
  relocate(Herd, .before = North)
  
w2.Q.sum <- w2.k2 %>% 
  group_by(Herd) %>% 
  summarise(
    QC1 = paste0(round(mean(North), 3), "±", round(sd(North), 3)),
    QC2 = paste0(round(mean(Central), 3), "±", round(sd(Central), 3))
  )

# Summary per herd of Q assignments
w2.Q.sum

write.csv(w2.Q.sum, "./output/tables/structure/WBNP.Cluster.assignment.csv")
```

### Combined plots 

```{r}
pcaw1_nl <- pcaw1 + 
  theme(
    plot.tag.position = "topleft",
    plot.tag = element_text(face = "bold", size = 25),
    plot.margin = margin(t = 0, r = 5, b = 0, l = 5, unit = "pt"),
    legend.position = "none"
  )

# pcaw2: Hide legend completely
pcaw2_nl <- pcaw2 + 
  guides(colour = guide_legend(ncol = 3, override.aes = list(size = 5))) +
  theme(
    plot.tag.position = "topleft",
    plot.tag = element_text(face = "bold", size = 25),
    plot.margin = margin(t = 0, r = 10, b = 10, l = 10, unit = "pt"),
  
    # Legend settings
    legend.position = c(0.98, 0.98),
    legend.justification = c("right", "top"),
    legend.background = element_rect(fill = alpha("white", 0.7), color = "grey80", size = 0.3),
    legend.key = element_blank(),
    legend.key.size = unit(0.35, "cm"),
    legend.title = element_text(face = "bold", size = 8),
    legend.text = element_text(size = 7),
    legend.margin = margin(2, 4, 2, 4, "pt"),
    legend.spacing.x = unit(0.05, "cm")
  )

wbnp_pcas <- grid.arrange(pcaw1_nl, pcaw2_nl, 
                            ncol = 2, 
                            widths = c(10, 10)
                            )
                            

# Save plot
library(svglite)
ggsave("./output/plots/PCAs/wbnp/wbnp_pcas.svg",
       plot = wbnp_pcas, 
       device = "svg", 
       width = 16, 
       height = 16, 
       units = "in")
# With plains bison
A_labeled <- arrangeGrob(admixture1$plot[[1]], 
                         top = textGrob("A", x = unit(0.02, "npc"), just = "left", 
                                        gp = gpar(fontsize = 25, fontface = "bold")))

B_labeled <- arrangeGrob(admixture2$plot[[1]], 
                         top = textGrob("B", x = unit(0.02, "npc"), just = "left",
                                        gp = gpar(fontsize = 25, fontface = "bold")))

# Combine Global no plains T1
top2 <- grid.arrange(A_labeled,B_labeled,
                      ncol = 1,
                      widths = 1,
                      padding = unit(0, "line"))

w_plains <- arrangeGrob(top2, admix_pcas, 
                        ncol = 1,
                        heights = c(15, 15),
                        widths = 1,
                        padding = unit(0, "line")
                        )

ggsave("./output/plots/admixedwbnp_combined.png",
      plot = w_plains,
       height = 15,
       width = 15,
       bg = "white",
       units = "in")

wbnp_cols <- struct_a1_cols <- c("#4477AA", "#EE6677", "#66CCEE", "#EE7733")


### Without plains

plotw1 <- plotQ(c(w1.mk2, w1.mk3, w1.mk4), 
      imgoutput = "join",
            basesize = 10, 
            grplab = loc_labels_w1, 
            subsetgrp = herd_order_w1, 
            clustercol = wbnp_cols,
            ordergrp = TRUE,
      
      # Side "K" Labels
            splab = c("K=2", "K=3", "K=4"),
            splabsize = 15,
            splabface = "bold",
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
      # Individual Label Settings
            showindlab = F, 
            indlabsize = 0,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 4.0, 
            grplabangle = 25,
            grplabpos = 0.75,
            grplabheight = 0.15,      
            grplabspacer = 0,
            grplabcol = "black",
      
            linesize = 1,          
            pointsize = 8,
            linepos = 0.95,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 1, 
            divtype = NULL,
            useindlab = TRUE, 
            returndata = TRUE,
            #barbordersize = 0.01,
            #barbordercolour = "lightgrey",
      
      # Add title
           showtitle = FALSE,
           titlelab = NULL,
           titlehjust = 0.5,
           titlevjust = 0.5,
           titlesize = 15,
           titlecol = "black",
           titleface = "bold",
      
      # Plot settings
            height = 10,             
            width = 35,
            panelspacer = 0.2,
            panelratio = c(1,2),
      returnplot = TRUE,
      
      # File settings
            imgtype = "tiff",
            outputfilename = "W1.k2-4",
            exportpath = "./output/plots/structure_plots/wbnp/")

plotw2 <- plotQ(c(w2.mk2, w2.mk3, w2.mk4), 
      imgoutput = "join",
            basesize = 10, 
            grplab = loc_labels_w2, 
            subsetgrp = herd_order_w2, 
            clustercol = wbnp_cols,
            ordergrp = TRUE,
      
      # Side "K" Labels
            splab = c("K=2", "K=3", "K=4"),
            splabsize = 15,
            splabface = "bold",
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
      # Individual Label Settings
            showindlab = F, 
            indlabsize = 0,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 4.0, 
            grplabangle = 25,
            grplabpos = 0.75,
            grplabheight = 0.15,      
            grplabspacer = 0,
            grplabcol = "black",
      
            linesize = 1,          
            pointsize = 8,
            linepos = 0.95,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 1, 
            divtype = NULL,
            useindlab = TRUE, 
            returndata = TRUE,
            #barbordersize = 0.01,
            #barbordercolour = "lightgrey",
      
      # Add title
           showtitle = FALSE,
           titlelab = NULL,
           titlehjust = 0.5,
           titlevjust = 0.5,
           titlesize = 15,
           titlecol = "black",
           titleface = "bold",
      
      # Plot settings
            height = 10,             
            width = 35,
            panelspacer = 0.2,
            panelratio = c(1,2),
      returnplot = TRUE,
      
      # File settings
            imgtype = "tiff",
            outputfilename = "W2.k2-4",
            exportpath = "./output/plots/structure_plots/wbnp/")


A_labeled <- arrangeGrob(plotw1$plot[[1]], 
                         top = textGrob("A", x = unit(0.2, "cm"), just = "left", 
                                        gp = gpar(fontsize = 25, fontface = "bold")))

B_labeled <- arrangeGrob(plotw2$plot[[1]], 
                         top = textGrob("B", x = unit(0.2, "cm"), just = "left",
                                        gp = gpar(fontsize = 25, fontface = "bold")))

C_labeled <- arrangeGrob(pcaw1_nl, 
                         top = textGrob("C", x = unit(0.2, "cm"), just = "left", 
                                         gp = gpar(fontsize = 25, fontface = "bold")))
D_labeled <- arrangeGrob(pcaw2_nl, 
                         top = textGrob("D", x = unit(0.2, "cm"), just = "left", 
                                         gp = gpar(fontsize = 25, fontface = "bold")))

wbnp_pcas <- grid.arrange(C_labeled, D_labeled, 
                            ncol = 2, 
                            widths = c(10, 10))
# Combine Global no plains T1
top3 <- grid.arrange(A_labeled,B_labeled,
                      ncol = 1,
                      widths = 1,
                      padding = unit(0, "line"))

# Shrink margins for PCA

no_plains <- arrangeGrob(top3, wbnp_pcas, 
                        ncol = 1,
                        heights = c(1, 0.8),
                        widths = 1,
                        padding = unit(-1, "line")
                        )

ggsave("./output/plots/combined_plots/wbnp_combined.png",
      plot = no_plains,
       height = 12,
       width = 12,
       bg = "white",
       units = "in")


```


```{r}
pcaa1_nl <- pcaa1 + 
  theme(
    plot.tag.position = "topleft",
    plot.tag = element_text(face = "bold", size = 25),
    plot.margin = margin(t =0, r = 10, b = 10, l = 10, unit = "pt"),
    legend.position = "none"
  )

pcaa2_nl <- pcaa2 + 
  guides(colour = guide_legend(ncol = 3, override.aes = list(size = 5))) +
  theme(
    plot.tag.position = "topleft",
    plot.tag = element_text(face = "bold", size = 25),
    plot.margin = margin(t = 0, r = 10, b = 10, l = 10, unit = "pt"),
  
    # Legend settings
    legend.position = c(0.98, 0.98),
    legend.justification = c("right", "top"),
    legend.background = element_rect(fill = alpha("white", 0.7), color = "grey80", size = 0.3),
    legend.key = element_blank(),
    legend.key.size = unit(0.35, "cm"),
    legend.title = element_text(face = "bold", size = 8),
    legend.text = element_text(size = 7),
    legend.margin = margin(2, 4, 2, 4, "pt"),
    legend.spacing.x = unit(0.05, "cm")
  )


custom_herd_order = c("EINPP", "WBNPLB", "WBNPNY", "WBNPHC", "WBNPGR", "WBNPPAD", "WENT", "EINPW")

# STRUCTURE T1 w plains
plot_a1k2 <- plotQ(a1.mk2,
            basesize = 10, 
            grplab = loc_labels_a1, 
            subsetgrp = custom_herd_order, 
            clustercol = master_cols,
            ordergrp = TRUE,
      
      # Side "K" Labels
            splab = "K=2",
            splabsize = 13,
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
            splabface = "bold",
      
            # Individual Label Settings
            showindlab = F, 
            indlabsize = 4,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 4, 
            grplabangle =25,
            grplabpos = 0.75,
            grplabheight = 0.2,      
            grplabspacer = 0.01,
            grplabcol = "black",
            linesize = 1,          
            pointsize = 8,
            linepos = 0.90,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 1, 
            divtype = NULL,
      
            useindlab = TRUE, 
            returndata = TRUE,
      
      # Add title
           showtitle = TRUE,
           titlelab = NULL,
           titlehjust = 0.5,
           titlevjust = 0.5,
           titlesize = 15,
           titlecol = "black",
      titleface = "bold",
      
      # Plot settings
            height = 10,             
            width = 25,
            panelspacer = 0,
            panelratio = c(1,2),
      
      # File settings
            imgtype = "tiff",
            returnplot = TRUE,
            outputfilename = "a1.k2",
            exportpath = "./output/plots/structure_plots/wbnp/")

custom_herd_order = c("EINPP", "WBNPLB", "WBNPNY", "WBNPHC", "WBNPGR", "WBNPPAD", "WENT", "EINPW")


# STRUCTURE T2 w plains
plot_a2k2 <- plotQ(a2.mk2,
            basesize = 15, 
            grplab = loc_labels_a2, 
            subsetgrp = custom_herd_order, 
            clustercol = master_cols,
            ordergrp = TRUE,

     # Side "K" Labels
            splab = "K=2",
            splabsize = 13,
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
            splabface = "bold",
      
            # Individual Label Settings
            showindlab = F, 
            indlabsize = 4,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 4, 
            grplabangle =25,
            grplabpos = 0.75,
            grplabheight = 0.2,      
            grplabspacer = 0.01,
            grplabcol = "black",
            linesize = 1,          
            pointsize = 8,
            linepos = 0.90,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 1, 
            divtype = NULL,
      
            useindlab = TRUE, 
            returndata = TRUE,
      
      # Add title
           showtitle = TRUE,
           titlelab = NULL,
           titlehjust = 0.5,
           titlevjust = 0.5,
           titlesize = 15,
           titlecol = "black",
      titleface = "bold",
      
      # Plot settings
            height = 10,             
            width = 25,
            panelspacer = 0,
            panelratio = c(1,2),
      # File settings
            imgtype = "tiff",
            returnplot = TRUE,
            outputfilename = "a2.k2",
            exportpath = "./output/plots/structure_plots/wbnp/")


# Combine plots w/ plains outgroup
library(gridExtra)

A_labeled <- arrangeGrob(plot_a1k2$plot[[1]], 
                         top = textGrob("A", x = unit(0.2, "cm"), just = "left", 
                                        gp = gpar(fontsize = 25, fontface = "bold")))

B_labeled <- arrangeGrob(plot_a2k2$plot[[1]], 
                         top = textGrob("B", x = unit(0.2, "cm"), just = "left",
                                        gp = gpar(fontsize = 25, fontface = "bold")))

C_labeled <- arrangeGrob(pcaa1_nl, 
                         top = textGrob("C", x = unit(0.2, "cm"), just = "left", 
                                         gp = gpar(fontsize = 25, fontface = "bold")))
D_labeled <- arrangeGrob(pcaa2_nl, 
                         top = textGrob("D", x = unit(0.2, "cm"), just = "left", 
                                         gp = gpar(fontsize = 25, fontface = "bold")))

global_pcas <- grid.arrange(C_labeled, D_labeled, 
                            ncol = 2, 
                            widths = c(10, 10))

# Save plot
library(svglite)
ggsave("./output/plots/PCAs/global/global_pcas.svg",
       plot = global_pcas, 
       device = "svg", 
       width = 16, 
       height = 12, 
       units = "in")

# Combine Global w/ plains T1
top1 <- arrangeGrob(
  A_labeled, 
  B_labeled,
  ncol = 1,
  heights = c(1, 1),
  padding = unit(-1.5, "lines") # Pulls plot B closer to plot A
)


wbnp_einp <- arrangeGrob(top1, global_pcas, 
                        ncol = 1,
                        heights = c(0.75, 1),
                        widths = 1,
                        padding = unit(-3, "lines")
                        )
# Save plots
ggsave("./output/plots/combined_plots/WBNP_EINP_outgroup.tiff",
       plot = wbnp_einp,
       height = 10,
       width = 12.5,
       bg = "white",
       units = "in")


```
```


