---
title: "WBNP_Complex_Structure"
author: "Alex Depew"
date: "2026-08-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# File information
This script is used to evaluate the presence and extent of genetic substructure within the Wood Buffalo National Park complex. 

Files used described below:

- `Final_dataset_2026.06.28.csv`
   - Contains all samples for population genetic analyses, post-filtering
   
## Load Required Packages
```{r}
rm(list = ls())
library(adegenet)
library(graph4lg)
library(poppr)
library(hierfstat)
library(pophelper)
library(tidyverse)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(ggpubr)
library(grid)
```

## Load Required Data
```{r}
cleaned.data <- read.csv("./datafiles/Final_dataset_2026.07.17.csv",
                         check.names = F)
```

#### Convert data to Genind object
```{r}
# Make genind object
bison.genind <- df2genind(X=cleaned.data[,c(15:ncol(cleaned.data))],
                              sep = "/",
                              ind.names = cleaned.data$Spreadsheet_ID,
                              loc.names = NULL,
                              pop = cleaned.data$Herd,
                              NA.char = "NA",
                              ploidy = 2,
                              type = "codom",
                              strata = NULL,
                              hierarchy = NULL)

# Attach collection year metadata
bison.genind@other$year <- cleaned.data$Year.Collected
```

#### Subset data 
Data is subset into two temporal groups. Temporal groups are: 1: pre-2016, and 3: post-2016
```{r}
# Create datasets containing samples of interest for temporal comparisons
c1 <- bison.genind[bison.genind@pop %in% c("WBNPGR", "WBNPHC", 
                                           "WBNPLB", "WBNPNY", 
                                           "WBNPPAD", "WENT", 
                                           "RL", "WAB", "MBS") &
                   bison.genind@other$year <=2015,]

c2 <- bison.genind[bison.genind@pop %in% c("WBNPGR", "WBNPHC", 
                                           "WBNPLB", "WBNPNY", 
                                           "WBNPPAD", "WENT", 
                                           "RL", "WAB", "MBS") &
                   bison.genind@other$year >=2016,]
```

# Principal Coordinate Analysis (PCA)
Visualize population structure without assumptions of HWE or LD, to identify clusters in the data.
```{r}
herds <- c( "WBNPLB", "WBNPNY", "WBNPHC", "WBNPPAD","WBNPGR", "WENT", "RL", "WAB", "MBS")

herd_palette <- c("#CCBB44","#228833", "#4477AA", "#EE6677",  "#66CCEE", "#AA3377", "#BBBBBB", "#EE7733", "#332288")

herd_cols <- setNames(as.character(herd_palette), herds)

```

# PCA
### Perform PCA on both datasets
```{r}
# Perform PCA
x.c1 <- tab(c1, freq = TRUE, NA.method = "mean")
pca.c1 <- dudi.pca(x.c1, center = T, scale = F, scannf = F, nf = 2)

# Axes % Variance
eig.perc.c1 <- 100*pca.c1$eig/sum(pca.c1$eig)

x.c2 <- tab(c2, freq = TRUE, NA.method = "mean")
pca.c2 <- dudi.pca(x.c2, center = T, scale = F, scannf = F, nf = 2)

# Axes % Variance
eig.perc.c2 <- 100*pca.c2$eig/sum(pca.c2$eig)

#### axis limits
max_x <- max(abs(c(pca.c1$li[, 1], pca.c2$li[, 1]))) * 1.10
max_y <- max(abs(c(pca.c1$li[, 2], pca.c2$li[, 2]))) * 1.10

xlims <- c(-max_x, max_x)
ylims <- c(-max_y, max_y)
```

## PCA plot with ggplot
#### WBNP Complex Temp1
```{r}
c1@pop <- factor(c1@pop, levels = sort(unique(as.character(c1@pop))))
levels_c1 <- levels(c1@pop)
c1_cols <- herd_cols[levels_c1]

pca.c1$li[ ,1] <- pca.c1$li[ ,1] * -1

# Plot PCA object
pcac1 <- fviz_pca_ind(pca.c1,
             axes = c(1, 2),
             geom = "blank",
             habillage = c1@pop,
             repel = TRUE,
             palette = c1_cols,
             alpha.ind = 1,) + 
  stat_ellipse(aes(color = c1@pop), type = "norm", level = 0.68) + 
  geom_point(aes(colour = c1@pop),
             shape = 19,
             alpha = 0.75,
             size = 2) + 
coord_fixed(ratio = 1, xlim = c(-1.5, 3), ylim = c(-2, 2)) +
  
theme(
    plot.title = element_blank(),
    legend.title = element_text(face = "bold", size = 15),
    legend.text = element_text(size = 15),
    axis.title = element_text(face = "bold", size = 20),
    text = element_text(size = 20),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),  ) + 


  labs(title = NULL,
       # Fixed the placement of closing parentheses and commas below
       x = paste0("PC1 (", round(eig.perc.c1[1], 2), "%)"),
       y = paste0("PC2 (", round(eig.perc.c1[2], 2), "%)"),
       colour = "Herd") + 
    guides(colour = guide_legend(override.aes = list(size = 4, shape = 19)))

       

pcac1

ggsave("./output/plots/PCAs/wbnp_complex//temp1.svg",
       plot = pcac1, 
       device = "svg", 
       width = 12, 
       height = 12, 
       units = "in")
```
#### WBNP Complex Temp2
```{r}
c2@pop <- factor(c2@pop, levels = sort(unique(as.character(c2@pop))))
levels_c2 <- levels(c2@pop)
c2_cols <- herd_cols[levels_c2]

#pca.c2$li[ ,1] <- pca.c2$li[ ,1] * -1
pca.c2$li[ ,2] <- pca.c2$li[ ,2] * -1


# Plot PCA object
pcac2 <- fviz_pca_ind(pca.c2,
             axes = c(1, 2),
             geom = "blank",
             habillage = c2@pop,
             repel = TRUE,
             palette = c2_cols,
             alpha.ind = 1,) + 
  stat_ellipse(aes(color = c2@pop), type = "norm", level = 0.68) + 
  geom_point(aes(colour = c2@pop),
             shape = 19,
             alpha = 0.75,
             size = 2) + 
#coord_fixed(ratio = 1, xlim = c(-1.5, 2.75), ylim = c(-1.8, 3.2)) +
  
theme(
    plot.title = element_blank(),
    legend.title = element_text(face = "bold", size = 15),
    legend.text = element_text(size = 15),
    axis.title = element_text(face = "bold", size = 20),
    text = element_text(size = 20),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),  ) + 


  labs(title = NULL,
       # Fixed the placement of closing parentheses and commas below
       x = paste0("PC1 (", round(eig.perc.c2[1], 2), "%)"),
       y = paste0("PC2 (", round(eig.perc.c2[2], 2), "%)"),
       colour = "Herd") + 
    guides(colour = guide_legend(override.aes = list(size = 4, shape = 19)))

       

pcac2

ggsave("./output/plots/PCAs/wbnp_complex//temp2.svg",
       plot = pcac2, 
       device = "svg", 
       width = 12, 
       height = 12, 
       units = "in")
```
# Create STRUCTURE Files
```{r}
summary(c1)
# W1 structure (202 individuals, 9 pops)
c1.str <- c1 
pop(c1.str) <- as.numeric(as.factor(pop(c1.str)))

genind_to_structure(c1.str,
                    output = "STRUCTURE/wbnp_complex/c1.structure.txt")

summary(c2)
# W2 structure (269 individuals, 9 pops)
c2.str <- c2
pop(c2.str) <- as.numeric(as.factor(pop(c2.str)))

genind_to_structure(c2.str,
                    output = "STRUCTURE/wbnp_complex/c2.structure.txt")
```
# Load Structure Files
```{r}
# List files
# Full datasets
sfiles_c1 <- list.files("./STRUCTURE/wbnp_complex/results/temp1/", full.names = TRUE)
sfiles_c2 <- list.files("./STRUCTURE/wbnp_complex/results/temp2/", full.names = TRUE)

# Qlist (pophelper)
slist_c1 <- readQ(files = sfiles_c1, filetype = "structure")
slist_c2 <- readQ(files = sfiles_c2, filetype = "structure")

# Summarise runs
struct_sum_c1 <- summariseQ(tabulateQ(slist_c1))
struct_sum_c2 <- summariseQ(tabulateQ(slist_c2))

### Likelihood plots
#### C1
c1.k <- evannoMethodStructure(data = struct_sum_c1, exportplot = F,
                           returnplot = T, returndata = F,
                           basesize = 25, linesize = 0.7,
                           height = 25, width = 25, units = "cm",
                           outputfilename = "c1_K",
                           exportpath = "./STRUCTURE_files/",
                           imgtype = "pdf")

p_lik <- c1.k$grobs[[1]]
p_dk <- c1.k$grobs[[4]]


# Label Likelihood as 'a'
panel_a <- arrangeGrob(p_lik, top = textGrob("A", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

# Label Delta K as 'b'
panel_b <- arrangeGrob(p_dk, top = textGrob("B", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

evanno_c1 <- grid.arrange(panel_a, panel_b, 
                         ncol = 2, 
                         widths = c(1, 1))

ggsave("./output/plots/K_plots/c1.k.svg", 
       plot = evanno_c1, 
       width = 27, 
       height = 11)

# WBNP2 ----
# K plots
c2.k <- evannoMethodStructure(data = struct_sum_c2, exportplot = F,
                           returnplot = T, returndata = F,
                           basesize = 25, linesize = 0.7,
                           height = 25, width = 25, units = "cm",
                           outputfilename = "c2_K",
                           exportpath = "./STRUCTURE_files/",
                           imgtype = "pdf")

p_lik <- c2.k$grobs[[1]]
p_dk <- c2.k$grobs[[4]]


# Label Likelihood as 'a'
panel_a <- arrangeGrob(p_lik, top = textGrob("A", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

# Label Delta K as 'b'
panel_b <- arrangeGrob(p_dk, top = textGrob("B", x = unit(0.5, "line"), y = unit(0.8, "npc"), 
                                             just = c("left", "top"), 
                                             gp = gpar(fontsize = 35, fontface = "bold")))

evanno_c2 <- grid.arrange(panel_a, panel_b, 
                         ncol = 2, 
                         widths = c(1, 1))

ggsave("./output/plots/K_plots/c2.k.svg", 
       plot = evanno_c2, 
       width = 27, 
       height = 11)

```

### Structure Plots
```{r}
# C1
c1.data <- cleaned.data %>% 
  filter(Year.Collected <=2015) %>%
  rename(ID = Spreadsheet.ID) %>% 
  mutate(ID = as.character(ID)) %>% 
  filter(Herd %in% c("WBNPGR", "WBNPHC", 
                     "WBNPLB", "WBNPNY", 
                     "WBNPPAD", "WENT", 
                     "RL", "WAB", "MBS"))
# Align runs
slist_c1 <- alignK(slist_c1, "auto")

# Merge runs (average over 10 replicates)
c1.mk2 <- mergeQ(slist_c1[11:20])
c1.mk3 <- mergeQ(slist_c1[21:30])
c1.mk4 <- mergeQ(slist_c1[31:40])
c1.mk5 <- mergeQ(slist_c1[41:50])
c1.mk6 <- mergeQ(slist_c1[51:60])
c1.mk7 <- mergeQ(slist_c1[61:70])
c1.mk8 <- mergeQ(slist_c1[71:80])
c1.mk9 <- mergeQ(slist_c1[81:90])


c1.merged.list <- as.qlist(list(c1.mk2[[1]], c1.mk3[[1]], 
                                c1.mk4[[1]], c1.mk5[[1]], 
                                c1.mk6[[1]], c1.mk7[[1]], 
                                c1.mk8[[1]], c1.mk9[[1]]))

ind_label_c1 <- readQ(files = sfiles_c1[1], indlabfrom = TRUE) %>% as.data.frame()
ind_label_c1 <- rownames_to_column(ind_label_c1, "ID")
locations_c1 <- left_join(ind_label_c1, c1.data, by = "ID")

loc_labels_c1 <- locations_c1$Herd %>% as.data.frame()

# C2
c2.data <- cleaned.data %>% 
  filter(Year.Collected >=2016) %>%
  rename(ID = Spreadsheet.ID) %>% 
  mutate(ID = as.character(ID)) %>% 
  filter(Herd %in% c("WBNPGR", "WBNPHC", 
                     "WBNPLB", "WBNPNY", 
                     "WBNPPAD", "WENT", 
                     "RL", "WAB", "MBS"))
# Align runs
slist_c2 <- alignK(slist_c2, "auto")

# Merge runs (average over 10 replicates)
c2.mk2 <- mergeQ(slist_c2[11:20])
c2.mk3 <- mergeQ(slist_c2[21:30])
c2.mk4 <- mergeQ(slist_c2[31:40])
c2.mk5 <- mergeQ(slist_c2[41:50])
c2.mk6 <- mergeQ(slist_c2[51:60])
c2.mk7 <- mergeQ(slist_c2[61:70])
c2.mk8 <- mergeQ(slist_c2[71:80])
c2.mk9 <- mergeQ(slist_c2[81:90])


c2.merged.list <- as.qlist(list(c2.mk2[[1]], c2.mk3[[1]], 
                                c2.mk4[[1]], c2.mk5[[1]], 
                                c2.mk6[[1]], c2.mk7[[1]], 
                                c2.mk8[[1]], c2.mk9[[1]]))

ind_label_c2 <- readQ(files = sfiles_c2[1], indlabfrom = TRUE) %>% as.data.frame()
ind_label_c2 <- rownames_to_column(ind_label_c2, "ID")
locations_c2 <- left_join(ind_label_c2, c2.data, by = "ID")

loc_labels_c2 <- locations_c2$Herd %>% as.data.frame()


# Set herd order and colour palette
herd_order <- c("RL", "WAB", "WBNPLB", "WBNPNY", "WBNPHC", "WBNPPAD", "WBNPGR","WENT", "MBS")
struct_cols <- c("#CCBB44","#228833","#66CCEE",  "#EE6677",  "#332288", "#EE7733","#4477AA","#AA3377", "#BBBBBB")
```

## C1
```{r}
# Merge runs (average over 10 replicates)
c1.mk2 <- mergeQ(slist_c1[11:20])
c1.mk3 <- mergeQ(slist_c1[21:30])
c1.mk4 <- mergeQ(slist_c1[31:40])
c1.mk5 <- mergeQ(slist_c1[41:50])
c1.mk6 <- mergeQ(slist_c1[51:60])
c1.mk7 <- mergeQ(slist_c1[61:70])
c1.mk8 <- mergeQ(slist_c1[71:80])
c1.mk9 <- mergeQ(slist_c1[81:90])

#mk2
Cluster2 <- c1.mk2$`2`$Cluster1 
Cluster1 <- c1.mk2$`2`$Cluster2 

c1.mk2$`2`$Cluster1 <- Cluster1
c1.mk2$`2`$Cluster2 <- Cluster2

#mk3
Cluster3 <- c1.mk3$`3`$Cluster1
Cluster2 <- c1.mk3$`3`$Cluster2 
Cluster1 <- c1.mk3$`3`$Cluster3 

c1.mk3$`3`$Cluster1 <- Cluster1
c1.mk3$`3`$Cluster2 <- Cluster2
c1.mk3$`3`$Cluster3 <- Cluster3

#mk4
Cluster1 <- c1.mk4$`4`$Cluster1
Cluster4 <- c1.mk4$`4`$Cluster2 
Cluster3 <- c1.mk4$`4`$Cluster3 
Cluster2 <- c1.mk4$`4`$Cluster4 

c1.mk4$`4`$Cluster1 <- Cluster1
c1.mk4$`4`$Cluster2 <- Cluster2
c1.mk4$`4`$Cluster3 <- Cluster3
c1.mk4$`4`$Cluster4 <- Cluster4

#mk5
Cluster1 <- c1.mk5$`5`$Cluster1
Cluster4 <- c1.mk5$`5`$Cluster2 
Cluster3 <- c1.mk5$`5`$Cluster3 
Cluster2 <- c1.mk5$`5`$Cluster4 
Cluster5 <- c1.mk5$`5`$Cluster5 

c1.mk5$`5`$Cluster1 <- Cluster1
c1.mk5$`5`$Cluster2 <- Cluster2
c1.mk5$`5`$Cluster3 <- Cluster3
c1.mk5$`5`$Cluster4 <- Cluster4
c1.mk5$`5`$Cluster5 <- Cluster5

#mk6
Cluster6 <- c1.mk6$`6`$Cluster1
Cluster4 <- c1.mk6$`6`$Cluster2 
Cluster3 <- c1.mk6$`6`$Cluster3 
Cluster2 <- c1.mk6$`6`$Cluster4 
Cluster5 <- c1.mk6$`6`$Cluster5 
Cluster1 <- c1.mk6$`6`$Cluster6 

c1.mk6$`6`$Cluster1 <- Cluster1
c1.mk6$`6`$Cluster2 <- Cluster2
c1.mk6$`6`$Cluster3 <- Cluster3
c1.mk6$`6`$Cluster4 <- Cluster4
c1.mk6$`6`$Cluster5 <- Cluster5
c1.mk6$`6`$Cluster6 <- Cluster6


plotQ(c(c1.mk2, c1.mk3, c1.mk4, c1.mk5, c1.mk6), 
      imgoutput = "join",
            basesize = 10, 
            grplab = loc_labels_c1, 
            subsetgrp = herd_order, 
            clustercol = struct_cols,
            ordergrp = TRUE,
      
      # Side "K" Labels
            splab = c("K = 2", "K = 3", "K = 4", "K = 5", "K = 6"),
            splabsize = 35,
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
            # Individual Label Settings
            showindlab = F, 
            indlabsize = 4,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 10, 
            grplabangle =45,
            grplabpos = 0.65,
            grplabheight = 1,      
            grplabspacer = 0.1,
            grplabcol = "black",
      
            linesize = 2,          
            pointsize = 15,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 3, 
            divtype = NULL,
      
            useindlab = TRUE, 
            returndata = TRUE,
      
      # Plot settings
            height = 7,             
            width = 50,
            panelspacer = 1,
            panelratio = c(1,1),
      
      
      # File settings
            imgtype = "tiff",
            outputfilename = "C1",
            exportpath = "./output/plots/structure_plots/wbnp_complex")
```

## C2
```{r}
# Merge runs (average over 10 replicates)
c2.mk2 <- mergeQ(slist_c2[11:20])
c2.mk3 <- mergeQ(slist_c2[21:30])
c2.mk4 <- mergeQ(slist_c2[31:40])
c2.mk5 <- mergeQ(slist_c2[41:50])
c2.mk6 <- mergeQ(slist_c2[51:60])
c2.mk7 <- mergeQ(slist_c2[61:70])
c2.mk8 <- mergeQ(slist_c2[71:80])
c2.mk9 <- mergeQ(slist_c2[81:90])

#mk2
Cluster2 <- c2.mk2$`2`$Cluster1 
Cluster1 <- c2.mk2$`2`$Cluster2 

c2.mk2$`2`$Cluster1 <- Cluster1
c2.mk2$`2`$Cluster2 <- Cluster2

#mk3
Cluster2 <- c2.mk3$`3`$Cluster1
Cluster3 <- c2.mk3$`3`$Cluster2 
Cluster1 <- c2.mk3$`3`$Cluster3 

c2.mk3$`3`$Cluster1 <- Cluster1
c2.mk3$`3`$Cluster2 <- Cluster2
c2.mk3$`3`$Cluster3 <- Cluster3

#mk4
Cluster3 <- c2.mk4$`4`$Cluster1
Cluster2 <- c2.mk4$`4`$Cluster2 
Cluster4 <- c2.mk4$`4`$Cluster3 
Cluster1 <- c2.mk4$`4`$Cluster4 

c2.mk4$`4`$Cluster1 <- Cluster1
c2.mk4$`4`$Cluster2 <- Cluster2
c2.mk4$`4`$Cluster3 <- Cluster3
c2.mk4$`4`$Cluster4 <- Cluster4

#mk5
Cluster4 <- c2.mk5$`5`$Cluster1
Cluster2 <- c2.mk5$`5`$Cluster2 
Cluster5 <- c2.mk5$`5`$Cluster3 
Cluster1 <- c2.mk5$`5`$Cluster4 
Cluster3 <- c2.mk5$`5`$Cluster5 

c2.mk5$`5`$Cluster1 <- Cluster1
c2.mk5$`5`$Cluster2 <- Cluster2
c2.mk5$`5`$Cluster3 <- Cluster3
c2.mk5$`5`$Cluster4 <- Cluster4
c2.mk5$`5`$Cluster5 <- Cluster5

#mk6
Cluster4 <- c2.mk6$`6`$Cluster1
Cluster1 <- c2.mk6$`6`$Cluster2 
Cluster2 <- c2.mk6$`6`$Cluster3 
Cluster3 <- c2.mk6$`6`$Cluster4 
Cluster5 <- c2.mk6$`6`$Cluster5 
Cluster6 <- c2.mk6$`6`$Cluster6 

c2.mk6$`6`$Cluster1 <- Cluster1
c2.mk6$`6`$Cluster2 <- Cluster2
c2.mk6$`6`$Cluster3 <- Cluster3
c2.mk6$`6`$Cluster4 <- Cluster4
c2.mk6$`6`$Cluster5 <- Cluster5
c2.mk6$`6`$Cluster6 <- Cluster6


plotQ(c(c2.mk2, c2.mk3, c2.mk4, c2.mk5, c2.mk6), 
      imgoutput = "join",
            basesize = 10, 
            grplab = loc_labels_c2, 
            subsetgrp = herd_order, 
            clustercol = struct_cols,
            ordergrp = TRUE,
      
      # Side "K" Labels
            splab = c("K = 2", "K = 3", "K = 4", "K = 5", "K = 6"),
            splabsize = 35,
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
            # Individual Label Settings
            showindlab = F, 
            indlabsize = 4,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 10, 
            grplabangle =45,
            grplabpos = 0.65,
            grplabheight = 1,      
            grplabspacer = 0.1,
            grplabcol = "black",
      
            linesize = 2,          
            pointsize = 15,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 3, 
            divtype = NULL,
      
            useindlab = TRUE, 
            returndata = TRUE,
      
      # Plot settings
            height = 7,             
            width = 50,
            panelspacer = 1,
            panelratio = c(1,1),
      
      
      # File settings
            imgtype = "tiff",
            outputfilename = "C2",
            exportpath = "./output/plots/structure_plots/wbnp_complex")
```


### Combined PCA and Structure Plots
```{r}
pcac1_nl <- pcac1 + 
  theme(
    plot.tag.position = "topleft",
    plot.tag = element_text(face = "bold", size = 25),
    plot.margin = margin(t = 0, r = 5, b = 0, l = 5, unit = "pt"),
    legend.position = "none"
  )

pcac2_nl <- pcac2 + 
  guides(colour = guide_legend(ncol = 3, override.aes = list(size = 6.5))) +
  theme(
    plot.tag.position = "topleft",
    plot.tag = element_text(face = "bold", size = 25),
    plot.margin = margin(t = 0, r = 10, b = 10, l = 10, unit = "pt"),
  
    # Legend settings
    legend.position = c(0.02, 0.98),
    legend.justification = c("left", "top"),
    legend.background = element_rect(fill = alpha("white", 0.7), color = "grey80", size = 0.3),
    legend.key = element_blank(),
    legend.key.size = unit(0.35, "cm"),
    legend.title = element_text(face = "bold", size = 8),
    legend.text = element_text(size = 7),
    legend.margin = margin(2, 4, 2, 4, "pt"),
    legend.spacing.x = unit(0.05, "cm")
    
  )


wbnp_complex_pcas <- grid.arrange(pcac1_nl, pcac2_nl, 
                            ncol = 2, 
                            widths = c(10, 10)
                            )
                            

# Save plot
library(svglite)
ggsave("./output/plots/PCAs/wbnp_complex/wbnp_complex_pcas.svg",
       plot = wbnp_complex_pcas, 
       device = "svg", 
       width = 15, 
       height = 7, 
       units = "in")


# C1 Temp1
plotc1 <- plotQ(c(c1.mk2, c1.mk3, c1.mk4, c1.mk5), 
      imgoutput = "join",
            basesize = 10, 
            grplab = loc_labels_c1, 
            subsetgrp = herd_order, 
            clustercol = struct_cols,
            ordergrp = TRUE,
      
      # Side "K" Labels
            splab = c("K=2", "K=3", "K=4", "K=5"),
            splabsize = 15,
            splabface = "bold",
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
      # Individual Label Settings
            showindlab = F, 
            indlabsize = 0,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 4.0, 
            grplabangle = 25,
            grplabpos = 0.75,
            grplabheight = 0.15,      
            grplabspacer = 0,
            grplabcol = "black",
      
            linesize = 1,          
            pointsize = 8,
            linepos = 0.95,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 1, 
            divtype = NULL,
            useindlab = TRUE, 
            returndata = TRUE,
            #barbordersize = 0.01,
            #barbordercolour = "lightgrey",
      
      # Add title
           showtitle = FALSE,
           titlelab = NULL,
           titlehjust = 0.5,
           titlevjust = 0.5,
           titlesize = 15,
           titlecol = "black",
           titleface = "bold",
      
      # Plot settings
            height = 10,             
            width = 35,
            panelspacer = 0.2,
            panelratio = c(1,2),
      returnplot = TRUE,
      
      # File settings
            imgtype = "tiff",
            outputfilename = "C1.k2-5",
            exportpath = "./output/plots/structure_plots/wbnp_complex//")

plotc2 <- plotQ(c(c2.mk2, c2.mk3, c2.mk4, c2.mk5), 
      imgoutput = "join",
            basesize = 10, 
            grplab = loc_labels_c2, 
            subsetgrp = herd_order, 
            clustercol = struct_cols,
            ordergrp = TRUE,
      
      # Side "K" Labels
            splab = c("K=2", "K=3", "K=4", "K=5"),
            splabsize = 15,
            splabface = "bold",
            sppos = "left",
            splabcol = "black",
            splabangle = 0,
      
      # Individual Label Settings
            showindlab = F, 
            indlabsize = 0,        
            indlabangle = 90,
            indlabheight = 0,
      
      # Group Label settings
            grplabsize = 4.0, 
            grplabangle = 25,
            grplabpos = 0.75,
            grplabheight = 0.15,      
            grplabspacer = 0,
            grplabcol = "black",
      
            linesize = 1,          
            pointsize = 8,
            linepos = 0.95,
      
      # Divider Line Settings
            divcol = "white",
            divsize = 1, 
            divtype = NULL,
            useindlab = TRUE, 
            returndata = TRUE,
            #barbordersize = 0.01,
            #barbordercolour = "lightgrey",
      
      # Add title
           showtitle = FALSE,
           titlelab = NULL,
           titlehjust = 0.5,
           titlevjust = 0.5,
           titlesize = 15,
           titlecol = "black",
           titleface = "bold",
      
      # Plot settings
            height = 10,             
            width = 35,
            panelspacer = 0.2,
            panelratio = c(1,2),
      returnplot = TRUE,
      
      # File settings
            imgtype = "tiff",
            outputfilename = "C2.k2-5",
            exportpath = "./output/plots/structure_plots/wbnp_complex//")


A_labeled <- arrangeGrob(plotc1$plot[[1]], 
                         top = textGrob("A", x = unit(0.2, "cm"), just = "left", 
                                        gp = gpar(fontsize = 25, fontface = "bold")))

B_labeled <- arrangeGrob(plotc2$plot[[1]], 
                         top = textGrob("B", x = unit(0.2, "cm"), just = "left",
                                        gp = gpar(fontsize = 25, fontface = "bold")))

C_labeled <- arrangeGrob(pcac1_nl, 
                         top = textGrob("C", x = unit(0.2, "cm"), just = "left", 
                                         gp = gpar(fontsize = 25, fontface = "bold")))
D_labeled <- arrangeGrob(pcac2_nl, 
                         top = textGrob("D", x = unit(0.2, "cm"), just = "left", 
                                         gp = gpar(fontsize = 25, fontface = "bold")))

wbnp_complex_pcas <- grid.arrange(C_labeled, D_labeled, 
                            ncol = 2, 
                            widths = c(10, 10))
# Combine Global no plains T1
top3 <- grid.arrange(A_labeled,B_labeled,
                      ncol = 1,
                      widths = 1,
                      padding = unit(0, "line"))

# Shrink margins for PCA

wbnp_complex <- arrangeGrob(top3, wbnp_complex_pcas, 
                        ncol = 1,
                        heights = c(1, 0.8),
                        widths = 1,
                        padding = unit(-1, "line")
                        )

ggsave("./output/plots/combined_plots/wbnp_complex_combined.png",
      plot = wbnp_complex,
       height = 12,
       width = 12,
       bg = "white",
       units = "in")


```
